---
title: "RTSL Cardiovascular Health Program Impact Assessment: Projected Health Outcomes 2017–2050"
date: "`r Sys.Date()`"
author: "University of Washington"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
    theme: flatly
    highlight: tango
    code_folding: hide
---

```{r setup, include=FALSE, echo=FALSE}
rm(list = ls())
knitr::opts_chunk$set(
  echo    = TRUE,
  error   = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.height = 5,
  fig.width  = 9,
  fig.align  = "center"
)
```

```{r libraries, message=FALSE, warning=FALSE, echo=FALSE}
library(data.table)
library(ggplot2)
library(knitr)
library(kableExtra)
library(DT)
library(scales)
library(openxlsx)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
```

```{r paths, message=FALSE, warning=FALSE, echo=FALSE}
wd       <- "C:/Users/wrgar/OneDrive - UW/02Work/ResolveToSaveLives/Assessment/uw-rtsl-assessment/"
wd_data  <- paste0(wd, "data/processed/")
wd_outp  <- paste0(wd, "output/")
wd_scen  <- paste0(wd, "scenarios/")
wd_raw   <- paste0(wd, "data/raw/")
```

```{r helpers, message=FALSE, warning=FALSE, echo=FALSE}
# ---- Formatters --------------------------------------------------------------
fmt_ci <- function(est, lo, hi, digits = 1) {
  f <- function(x) formatC(round(x / 1e6, digits), format = "f",
                            digits = digits, big.mark = ",")
  paste0(f(est), " (", f(lo), "\u2013", f(hi), ")")
}
fmt_m <- function(x, digits = 1)
  formatC(round(x / 1e6, digits), format = "f", digits = digits, big.mark = ",")
fmt_k <- function(x, digits = 0)
  formatC(round(x / 1e3, digits), format = "f", digits = digits, big.mark = ",")
fmt_M <- function(x, digits = 1)
  formatC(round(x / 1e6, digits), format = "f", digits = digits, big.mark = ",")

# ---- Cause labels ------------------------------------------------------------
cause_map <- c(
  ihd     = "Ischemic heart disease",
  istroke = "Ischemic stroke",
  hstroke = "Intracerebral hemorrhage",
  hhd     = "Hypertensive heart disease"
)

# ---- Intervention display labels for RTSL Assessment -------------------------
# Model output scenarios (from 06_run_scenarios_assessment.R):
#   baseline  -> Counterfactual (BAU)
#   bp_only   -> HTN Control (Program)
#   tfa_only  -> TFA Elimination
#   bp_tfa    -> HTN Control + TFA Elimination
#
# Each scenario is run twice via rate_scenario:
#   "program"   -> Observed RTSL program control rates (2017-2024, maintained to 2050)
#   "bau_upper" -> Counterfactual (zero incremental program effect)
#
# The program-attributable impact = deaths under bau_upper - deaths under program

int_labels <- c(
  "Counterfactual"                = "Counterfactual",
  "HTN Control (Program)"        = "HTN Control (Program)",
  "TFA Elimination"              = "TFA Elimination",
  "HTN Control + TFA Elimination" = "HTN Control + TFA Elimination"
)

int_order <- c(
  "HTN Control (Program)",
  "TFA Elimination",
  "HTN Control + TFA Elimination"
)

# ---- Colour palettes ---------------------------------------------------------
int_cols <- c(
  "HTN Control (Program)"         = "#1f78b4",
  "TFA Elimination"               = "#e31a1c",
  "HTN Control + TFA Elimination" = "#6a3d9a",
  "Counterfactual"                = "#999999"
)

cause_cols <- c(
  ihd     = "#56B4E9",
  istroke = "#CC79A7",
  hstroke = "#F0E442",
  hhd     = "#E69F00"
)

cause_labels_plot <- c(
  ihd     = "Ischemic heart disease",
  istroke = "Ischemic stroke",
  hstroke = "Intracerebral hemorrhage",
  hhd     = "Hypertensive heart disease"
)

# ---- WHO standard population (ages 20-95+) for ASMR -------------------------
full_who_std <- data.table(
  age    = 0:100,
  weight = c(
    rep(0.013, 10), rep(0.014, 10), rep(0.013, 10), rep(0.012, 10),
    rep(0.011, 10), rep(0.009, 10), rep(0.007, 10), rep(0.005, 10),
    rep(0.003, 10), rep(0.001, 10), 0.001
  )
)
who_std <- full_who_std[age >= 20 & age <= 94]
who_std <- rbind(
  who_std,
  data.table(age = 95, weight = full_who_std[age >= 95, sum(weight)])
)
who_std[, weight := weight / sum(weight)]

calculate_asmr <- function(dt, std_pop = who_std,
                            deaths_col = "deaths_dx",
                            population_col = "Nx",
                            disagg_var = "intervention") {
  dt <- copy(dt)
  dt[, age := as.integer(gsub("\\+", "", age))]
  dt <- merge(dt, std_pop, by = "age", all.x = TRUE)
  dt[, rate          := get(deaths_col) / get(population_col)]
  dt[, weighted_rate := rate * weight]
  group_vars <- c("year", disagg_var)
  dt[, .(ASMR = sum(weighted_rate, na.rm = TRUE) * 1e5), by = group_vars]
}

# ---- Deaths-delayed calculator -----------------------------------------------
deaths_summary <- function(dt,
                           baseline_label = "Counterfactual",
                           by_cols = c("age_group", "sex", "year",
                                       "location", "cause", "age"),
                           value_col = "dead") {
  stopifnot(is.data.table(dt))
  n_interventions <- uniqueN(dt[intervention != baseline_label, intervention])
  baseline_cols   <- setdiff(by_cols, "intervention")
  baseline   <- dt[intervention == baseline_label,
                   .(baseline_deaths = sum(get(value_col), na.rm = TRUE)),
                   by = baseline_cols]
  nonbaseline <- dt[intervention != baseline_label,
                    .(deaths_projected = sum(get(value_col), na.rm = TRUE)),
                    by = by_cols]
  if (!("intervention" %in% by_cols))
    nonbaseline[, deaths_projected := deaths_projected / n_interventions]
  merged <- merge(nonbaseline, baseline, by = baseline_cols,
                  all.x = TRUE, allow.cartesian = TRUE)
  merged[, deaths_delayed     := baseline_deaths - deaths_projected]
  merged[, deaths_delayed_ave := deaths_delayed / uniqueN(dt$year)]
  merged[, deaths_delayed_pct := deaths_delayed / baseline_deaths]
  merged[]
}
```

```{r country-metadata, message=FALSE, warning=FALSE, echo=FALSE}
countries_dt <- fread(paste0(wd_data, "Country_groupings_extended.csv"))

# ---- Standardise country names if needed ------------------------------------
countries_dt[location == "United States of America", location := "United States"]
countries_dt[location == "Congo, Dem. Rep.", location := "Democratic Republic of the Congo"]
countries_dt[location == "Vietnam", location := "Viet Nam"]
countries_dt[location == "Bolivia (Plurinational State of)", location := "Bolivia"]

# ---- Add missing regions for specific ISO3s ---------------------------------
new_countries <- data.table(
  iso3     = c("LCA", "VCT"),
  location = c("Saint Lucia", "Saint Vincent and the Grenadines"),
  region   = c("Latin America and Caribbean", "Latin America and Caribbean")
)
countries_dt <- rbind(countries_dt, new_countries[!iso3 %in% countries_dt$iso3],
                      use.names = TRUE, fill = TRUE)
```

```{r load-model-output, message=FALSE, warning=FALSE, echo=FALSE}
# ---- Load all country-level model output RDS files --------------------------
# Generated by 06_run_scenarios_assessment.R (program-impact mode)
out_model_path <- paste0(wd_outp, "out_model/")
files <- list.files(path = out_model_path, pattern = "\\.rds$", full.names = TRUE)

if (length(files) == 0)
  stop(paste0("No .rds files found in: ", out_model_path,
              "\nRun 06_run_scenarios_assessment.R first."))

data.out <- rbindlist(
  lapply(files, function(f) { dt <- readRDS(f); setDT(dt); dt }),
  use.names = TRUE, fill = TRUE
)
data.out <- data.out[!is.na(location)]

# Drop effect-ratio tracking columns if present
for (col in c("eff_ir", "eff_cf"))
  if (col %in% names(data.out)) data.out[, (col) := NULL]

# ---- Standardise scenario -> intervention labels ----------------------------
# The output has a 'scenario' column (baseline, bp_only, tfa_only, bp_tfa)
# and a 'rate_scenario' column (program, bau_upper)
if ("scenario" %in% names(data.out) && !"intervention" %in% names(data.out)) {
  setnames(data.out, "scenario", "intervention")
} else if ("scenario" %in% names(data.out)) {
  data.out[, intervention := scenario]
  data.out[, scenario := NULL]
}

# Ensure rate_scenario column exists
if (!"rate_scenario" %in% names(data.out))
  data.out[, rate_scenario := "program"]

# ---- Remap intervention labels -----------------------------------------------
data.out[intervention %in% c("baseline", "BAU", "b.a.u"),
         intervention := "Counterfactual"]
data.out[intervention %in% c("bp_only", "antihypertensive", "BP",
                              "Antihypertensive therapy"),
         intervention := "HTN Control (Program)"]
data.out[intervention %in% c("tfa_only", "tfa", "TFA"),
         intervention := "TFA Elimination"]
data.out[intervention %in% c("bp_tfa"),
         intervention := "HTN Control + TFA Elimination"]

# Keep only recognised scenarios
data.out <- data.out[intervention %in% c("Counterfactual", int_order)]

# ---- Separate program and counterfactual runs --------------------------------
# Program-attributable impact:
#   deaths_averted = deaths(bau_upper) - deaths(program)
# For the report, we use:
#   - rate_scenario == "bau_upper" as the Counterfactual
#   - rate_scenario == "program"   as the Intervention
data_program <- data.out[rate_scenario == "program"]
data_bau     <- data.out[rate_scenario == "bau_upper"]

# ---- Age groups --------------------------------------------------------------
data_program[, age_group := cut(age, breaks = c(0, 70, 80, 90, Inf),
                                labels = c("<70", "70-79", "80-89", "90+"),
                                right = FALSE, include.lowest = TRUE)]
data_bau[, age_group := cut(age, breaks = c(0, 70, 80, 90, Inf),
                            labels = c("<70", "70-79", "80-89", "90+"),
                            right = FALSE, include.lowest = TRUE)]

# ---- Merge region metadata ---------------------------------------------------
if (nrow(countries_dt) > 0) {
  region_lookup <- countries_dt[, .(location, region)]
  region_lookup <- unique(region_lookup)
  data_program <- merge(data_program, region_lookup, by = "location", all.x = TRUE)
  data_bau     <- merge(data_bau,     region_lookup, by = "location", all.x = TRUE)
}
if (!"region" %in% names(data_program)) data_program[, region := NA_character_]
if (!"region" %in% names(data_bau))     data_bau[, region := NA_character_]

# ---- Key summary counts ------------------------------------------------------
n_countries <- uniqueN(data_program$location)
year_min    <- min(data_program$year, na.rm = TRUE)
year_max    <- max(data_program$year, na.rm = TRUE)
study_years <- paste0(year_min, "\u2013", year_max)
```

---

# Background

This report presents modelled projections of cardiovascular disease (CVD) mortality
attributable to **Resolve to Save Lives (RTSL) cardiovascular health programs**
operating in `r n_countries` countries from 2017 to 2024. The central research question is:

> *What is the expected health impact --- in terms of deaths averted ---
> of RTSL cardiovascular health programs from 2017--2024 in targeted populations,
> projected to 2050?*

Two interventions are assessed:

1. **Improved hypertension control** --- observed program-level blood pressure control
   rates (2017--2024) are maintained at 2024 levels through 2050, compared to a
   counterfactual scenario with no incremental program effect.
2. **Industrial trans-fat elimination (TFA)** --- policy-driven elimination of
   industrial trans-fatty acids in countries with enacted or effective TFA policies
   by end of 2024.

Four causes of CVD death are modelled jointly: ischemic heart disease (IHD), ischemic
stroke, intracerebral hemorrhage (hemorrhagic stroke), and hypertensive heart disease (HHD).
A discrete-time multi-state transition model (Well $\rightarrow$ Sick $\rightarrow$ Dead) is calibrated to GBD 2023
data and projected forward to 2050 using UNWPP 2024 population estimates.

---

# Study Design

## Countries Included

The `r n_countries` countries with RTSL program data are:

```{r country-list, message=FALSE, warning=FALSE, echo=FALSE}
dt_countries <- unique(data_program[, .(location, region)])
setorder(dt_countries, region, location)
countries_by_region <- dt_countries[,
  .(Countries = paste0(sort(location), collapse = ", ")),
  by = .(Region = region)
]
setorder(countries_by_region, Region)

kable(countries_by_region,
      caption = "Countries included in the assessment, by region.",
      align = "ll") |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = TRUE, font_size = 13) |>
  row_spec(0, bold = TRUE)
```

## Intervention Definitions

| Intervention | Description |
|---|---|
| **HTN Control (Program)** | Antihypertensive therapy scale-up using observed RTSL program control rates. Baseline BP distributions are shifted based on Ettehad et al. cumulative effect sizes for each 10 mmHg BP bin. Program enrollment data (2017--2024) define the treated population; control rates are maintained at 2024 levels through 2050. |
| **TFA Elimination** | Complete elimination of industrial trans-fatty acids (target: 0% energy from TFA). Applied to countries with enacted/effective TFA best-practice policies by end of 2024, with a 2-year lag from policy passage to population-level effect. |
| **HTN Control + TFA Elimination** | Both interventions applied simultaneously with multiplicative effect combination. |

## Scenario Structure

The model is run in **program-rate mode**: each country is simulated twice:

1. **Program run** (`rate_scenario = "program"`): observed RTSL control rates applied
2. **Counterfactual run** (`rate_scenario = "bau_upper"`): zero incremental program effect

The **program-attributable impact** is the difference: deaths under counterfactual minus deaths under program.

## Projection Horizon

- **Program enrollment data**: 2017--2024 (from HTN global tracker)
- **Model projection**: `r study_years`
- **Impact assessment window**: 2017--2050
- **Age range**: 20--95 years

---

# Hypertension Control Program Data

```{r htn-program-data, message=FALSE, warning=FALSE, echo=FALSE, fig.width=10, fig.height=5}
htn_scenarios <- tryCatch(
  fread(paste0(wd_data, "htn_control_scenarios_by_loc_year.csv")),
  error = function(e) NULL
)

if (!is.null(htn_scenarios)) {
  # Plot observed + projected program control rates
  dt_plot <- htn_scenarios[year <= 2050, .(location, year,
                                            control_rate = control_rate_program,
                                            observed = control_rate_observed)]

  g <- ggplot(dt_plot, aes(x = year, y = control_rate * 100, group = location)) +
    geom_line(color = "grey60", linewidth = 0.6, alpha = 0.7) +
    # Highlight observed data points
    geom_point(data = dt_plot[observed == 1],
               aes(x = year, y = control_rate * 100),
               color = "#1f78b4", size = 0.8, alpha = 0.5) +
    scale_x_continuous(breaks = seq(2017, 2050, 5),
                       limits = c(2017, 2051), expand = c(0, 0)) +
    scale_y_continuous(limits = c(0, 60), expand = c(0, 0)) +
    labs(
      title    = "RTSL Hypertension Control Rates by Country",
      subtitle = "Observed (dots) and projected (maintained at 2024 level) program control rates",
      x = "Year", y = "Hypertension Control Rate (%)"
    ) +
    theme_minimal(base_size = 13) +
    theme(plot.title = element_text(face = "bold"), panel.grid.minor = element_blank())
  print(g)
  ggsave(paste0(wd_outp, "assessment_policy_htn_program.png"), plot = g,
         width = 11, height = 5, dpi = 300)
} else {
  message("htn_control_scenarios_by_loc_year.csv not found - skipping HTN program plot.")
}
```

```{r htn-enrollment, message=FALSE, warning=FALSE, echo=FALSE, fig.width=10, fig.height=5}
# # Program enrollment trajectory
# if (!is.null(htn_scenarios)) {
#   dt_enroll <- htn_scenarios[!is.na(enrolled) & enrolled > 0,
#                              .(location, year, enrolled)]
#   dt_enroll_total <- dt_enroll[, .(total_enrolled = sum(enrolled, na.rm = TRUE)),
#                                by = year]
# 
#   g <- ggplot() +
#     geom_line(data = dt_enroll,
#               aes(x = year, y = enrolled / 1e3, group = location),
#               color = "grey60", linewidth = 0.5, alpha = 0.6) +
#     geom_line(data = dt_enroll_total,
#               aes(x = year, y = total_enrolled / 1e6),
#               color = "#1f78b4", linewidth = 1.2) +
#     scale_x_continuous(breaks = seq(2017, 2025, 1)) +
#     scale_y_continuous(labels = comma) +
#     labs(
#       title    = "RTSL Hypertension Program Enrollment Over Time",
#       subtitle = "Grey = country-level (thousands); Blue = total across all countries (millions)",
#       x = "Year", y = "Enrolled (thousands / millions)"
#     ) +
#     theme_minimal(base_size = 13) +
#     theme(plot.title = element_text(face = "bold"), panel.grid.minor = element_blank())
#   print(g)
#   ggsave(paste0(wd_outp, "assessment_enrollment_trend.png"), plot = g,
#          width = 11, height = 5, dpi = 300)
# }
# Program enrollment trajectory (TOTAL only)
if (!is.null(htn_scenarios)) {

  dt_enroll <- htn_scenarios[
    !is.na(enrolled) & enrolled > 0,
    .(location, year, enrolled)
  ]

  dt_enroll_total <- dt_enroll[
    , .(total_enrolled = sum(enrolled, na.rm = TRUE)),
    by = year
  ]

  g <- ggplot(dt_enroll_total, aes(x = year, y = total_enrolled / 1e6)) +
    geom_line(color = "#1f78b4", linewidth = 1.3) +
    scale_x_continuous(breaks = seq(2015, 2050, 5)) +
    scale_y_continuous(labels = comma) +
    labs(
      title    = "RTSL Hypertension Program Enrollment Over Time",
      subtitle = "Total enrollment across all countries (millions)",
      x = "Year",
      y = "Enrolled (millions)"
    ) +
    theme_minimal(base_size = 13) +
    theme(
      plot.title = element_text(face = "bold"),
      panel.grid.minor = element_blank()
    )

  print(g)

  ggsave(
    paste0(wd_outp, "assessment_enrollment_trend.png"),
    plot = g,
    width = 11, height = 5, dpi = 300
  )
}
```

---

# Disease Trends (Baseline)

## Incidence and Prevalence Over Time

```{r disease-trends, message=FALSE, warning=FALSE, echo=FALSE, fig.width=10, fig.height=5}
# Use counterfactual (bau_upper) as the baseline trend
dt_trend <- data_bau[intervention == "Counterfactual" & year >= 2017,
                     .(new_cases = sum(newcases, na.rm = TRUE),
                       sick      = sum(sick,     na.rm = TRUE),
                       pop       = sum(pop,      na.rm = TRUE)),
                     by = year]

dt_melt <- melt(dt_trend, id.vars = "year",
                measure.vars  = c("new_cases", "sick"),
                variable.name = "metric", value.name = "count")
dt_melt[, metric_lbl := fifelse(metric == "new_cases", "Incident cases", "Prevalent cases")]

g <- ggplot(dt_melt, aes(x = year, y = count / 1e6,
                          color = metric_lbl, linetype = year > 2024)) +
  geom_line(linewidth = 1) +
  scale_color_manual(values = c("Incident cases"  = "steelblue",
                                "Prevalent cases" = "tomato"),
                     name = NULL) +
  scale_linetype_manual(values = c(`FALSE` = "solid", `TRUE` = "dashed"), guide = "none") +
  scale_y_continuous(labels = comma) +
  labs(
    title    = "Incident and Prevalent CVD Cases \u2014 Counterfactual Scenario",
    subtitle = paste0("All four causes combined across ", n_countries,
                      " RTSL countries; solid = historical, dashed = projected"),
    x = "Year", y = "Persons (millions)",
    caption = "Source: Model output; ages 20\u201395+"
  ) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "bottom", panel.grid.minor = element_blank(),
        plot.title = element_text(face = "bold"))
print(g)
ggsave(paste0(wd_outp, "assessment_disease_trend.png"), plot = g,
       width = 11, height = 5, dpi = 300)
```

## Age-Standardised Mortality Rate by Cause

```{r asmr-cause, message=FALSE, warning=FALSE, echo=FALSE, fig.width=11, fig.height=6}
dt_asmr_in <- data_bau[intervention == "Counterfactual" & year >= 2017,
                        .(deaths_dx = sum(dead, na.rm = TRUE),
                          well      = sum(well, na.rm = TRUE),
                          sick      = sum(sick, na.rm = TRUE)),
                        by = .(cause, age, year)]
dt_asmr_in[, Nx := well + sick]

asmr_cause <- calculate_asmr(dt_asmr_in, deaths_col = "deaths_dx",
                              population_col = "Nx", disagg_var = "cause")
asmr_base  <- asmr_cause[year == 2017, .(cause, ASMR_base = ASMR)]
asmr_cause <- merge(asmr_cause, asmr_base, by = "cause")
asmr_cause[, index := ASMR / ASMR_base * 100]
asmr_cause[, cause_lbl := fcase(
  cause == "ihd",     "Ischemic heart disease",
  cause == "istroke", "Ischemic stroke",
  cause == "hstroke", "Intracerebral hemorrhage",
  cause == "hhd",     "Hypertensive heart disease",
  default = cause
)]

g <- ggplot(asmr_cause, aes(x = year, y = index,
                             color = cause_lbl, linetype = year > 2024)) +
  geom_line(linewidth = 1) +
  geom_hline(yintercept = 100, linetype = "dotted", color = "grey50") +
  scale_color_manual(values = setNames(cause_cols, cause_labels_plot),
                     name = "Cause of death") +
  scale_linetype_manual(values = c(`FALSE` = "solid", `TRUE` = "dashed"), guide = "none") +
  labs(
    title    = "Age-Standardised CVD Mortality Rate Index \u2014 Counterfactual",
    subtitle = paste0("Index: 2017 = 100; ", n_countries, " RTSL countries"),
    x = "Year", y = "ASMR Index (2017 = 100)",
    caption = "WHO World Standard Population (2000\u20132025), rescaled to ages 20\u201395+"
  ) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "bottom", panel.grid.minor = element_blank(),
        plot.title = element_text(face = "bold"))
print(g)
ggsave(paste0(wd_outp, "assessment_asmr_cause.png"), plot = g,
       width = 11, height = 5, dpi = 300)
```

---

# Program-Attributable Deaths Averted (2017--2050)

```{r compute-deaths-averted, message=FALSE, warning=FALSE, echo=FALSE}
# ---- Core: program-attributable deaths averted --------------------------------
# deaths_averted = deaths(Counterfactual, bau_upper) - deaths(intervention, program)
#
# The counterfactual is ALWAYS "Counterfactual" in rate_scenario == "bau_upper".
# Merging on intervention was wrong for TFA: bau_upper has no "TFA Elimination" rows.

grp_keys <- c("age_group", "sex", "year", "location", "cause", "age")

# Counterfactual deaths: Counterfactual intervention, bau_upper rate scenario
dt_cf <- data_bau[intervention == "Counterfactual" & year >= 2017,
                  .(cf_deaths = sum(dead, na.rm = TRUE)),
                  by = grp_keys]

# Program deaths: all non-Counterfactual interventions, program rate scenario
dt_pgm_agg <- data_program[intervention %in% int_order & year >= 2017,
                            .(pgm_deaths = sum(dead, na.rm = TRUE)),
                            by = c(grp_keys, "intervention")]

# Merge: each intervention row gets the single counterfactual baseline
dt_merged <- merge(dt_pgm_agg, dt_cf, by = grp_keys, all.x = TRUE)
dt_merged[is.na(cf_deaths),  cf_deaths  := 0]
dt_merged[is.na(pgm_deaths), pgm_deaths := 0]
dt_merged[, deaths_averted := cf_deaths - pgm_deaths]

# ---- Cumulative totals per intervention --------------------------------------
dt_cumul <- dt_merged[, .(deaths_averted = sum(deaths_averted, na.rm = TRUE)),
                       by = intervention]

# ---- Annual time-series ------------------------------------------------------
dt_annual <- dt_merged[, .(deaths_averted = sum(deaths_averted, na.rm = TRUE)),
                        by = .(intervention, year)]
dt_annual[, intervention := factor(intervention, levels = int_order)]
setorder(dt_annual, intervention, year)
dt_annual[, cum_deaths_averted := cumsum(deaths_averted), by = intervention]

# ---- Per-scenario totals for inline text ------------------------------------
total_combined <- dt_cumul[intervention == "HTN Control + TFA Elimination", deaths_averted]
total_bp       <- dt_cumul[intervention == "HTN Control (Program)",         deaths_averted]
total_tfa      <- dt_cumul[intervention == "TFA Elimination",               deaths_averted]

n_years <- uniqueN(dt_merged$year)
```

```{r per1m-scalars, message=FALSE, warning=FALSE, echo=FALSE}
# Per-1M scalars for inline narrative (average per-country cumulative per 1M)
lives_per_1M_htn      <- round(total_bp       / n_countries)
lives_per_1M_tfa      <- round(total_tfa      / n_countries)
lives_per_1M_combined <- round(total_combined / n_countries)
```

Between 2017 and 2050, the combined RTSL hypertension control and trans-fat elimination
programs are projected to avert approximately **`r fmt_k(total_combined)` thousand deaths**
(`r fmt_k(lives_per_1M_combined)` thousand per million treated)
across `r n_countries` countries.

- **HTN Control (Program)** alone would avert `r fmt_k(total_bp)` thousand deaths
  (`r fmt_k(lives_per_1M_htn)` thousand per million treated).
- **TFA Elimination** alone would avert `r fmt_k(total_tfa)` thousand deaths
  (`r fmt_k(lives_per_1M_tfa)` thousand per million population).

## Table 1 --- Cumulative Deaths Averted by Intervention (2017--2050)

```{r table1, message=FALSE, warning=FALSE, echo=FALSE}
tab1 <- copy(dt_cumul)
tab1[, intervention := factor(intervention, levels = int_order)]
setorder(tab1, intervention)

tab1[, deaths_K      := round(deaths_averted / 1e3, 1)]
tab1[, deaths_annual := round(deaths_averted / n_years / 1e3, 1)]

tab1[, description := fcase(
  intervention == "HTN Control (Program)",
    "Observed RTSL program HTN control rates (2017\u20132024) maintained to 2050",
  intervention == "TFA Elimination",
    "Industrial TFA elimination via best-practice policies enacted by 2024",
  intervention == "HTN Control + TFA Elimination",
    "Both interventions combined (multiplicative effects)",
  default = ""
)]

tab1_fmt <- tab1[, .(
  Intervention       = intervention,
  Description        = description,
  Deaths_averted_K   = comma(deaths_K,      accuracy = 0.1),
  Avg_per_year       = comma(deaths_annual,  accuracy = 0.1)
)]
setnames(tab1_fmt,
         old = c("Description", "Deaths_averted_K", "Avg_per_year"),
         new = c("Intervention Description",
                 "Deaths averted (thousands, 2017\u20132050)",
                 "Average per year (thousands)"))

kable(tab1_fmt,
      caption = "Table 1. Cumulative program-attributable deaths averted (2017\u20132050) by intervention.",
      align   = "llrr") |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = TRUE, font_size = 13) |>
  column_spec(1, width = "6cm") |>
  column_spec(2, width = "10cm") |>
  row_spec(0, bold = TRUE) |>
  footnote(general = paste0(
    "Deaths averted = deaths under counterfactual (no program effect) minus deaths ",
    "under program scenario, summed over 2017\u20132050. ",
    "Causes: IHD, ischemic stroke, intracerebral hemorrhage, hypertensive heart disease."
  ))

dt_excel_tab1 <- copy(tab1_fmt)
```

## Figure 1 --- Deaths Averted by Intervention (Bar Chart)

```{r figure1, message=FALSE, warning=FALSE, echo=FALSE, fig.width=9, fig.height=5}
dt_fig1 <- tab1[, .(
  intervention   = factor(intervention, levels = rev(int_order)),
  deaths_averted = deaths_averted
)]

g <- ggplot(dt_fig1, aes(x = deaths_averted / 1e3, y = intervention,
                          fill = as.character(intervention))) +
  geom_col(width = 0.65) +
  scale_fill_manual(values = int_cols, guide = "none") +
  scale_x_continuous(expand = c(0, 0), labels = comma) +
  labs(
    title   = "Program-Attributable Deaths Averted by Intervention, 2017\u20132050",
    x       = "Deaths averted (thousands)",
    y       = NULL,
    caption = paste0("Deaths averted vs. counterfactual (no program effect); ",
                     "all CVD causes combined; ", n_countries, " RTSL countries.")
  ) +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold"),
        panel.grid.minor   = element_blank(),
        panel.grid.major.y = element_blank())
print(g)
ggsave(paste0(wd_outp, "assessment_fig1_deaths_intervention.png"), plot = g,
       width = 9, height = 5, dpi = 300)
```

## Table 2 --- Cumulative Deaths Averted by Intervention and Cause

```{r table2, message=FALSE, warning=FALSE, echo=FALSE}
dt_tab2 <- dt_merged[
  intervention %in% int_order,
  .(deaths_averted = sum(deaths_averted, na.rm = TRUE)),
  by = .(intervention, cause)
]

# TFA affects only IHD (via CHD risk pathway)
dt_tab2 <- dt_tab2[!(cause %in% c("istroke", "hstroke", "hhd") &
                        intervention == "TFA Elimination")]

dt_tab2[, cause_lbl := fcase(
  cause == "ihd",     "Ischemic heart disease",
  cause == "istroke", "Ischemic stroke",
  cause == "hstroke", "Intracerebral hemorrhage",
  cause == "hhd",     "Hypertensive heart disease",
  default = cause
)]

dt_tab2_wide <- dcast(dt_tab2, intervention ~ cause_lbl,
                      value.var = "deaths_averted", fill = NA)
dt_tab2_wide[, intervention := factor(intervention, levels = int_order)]
setorder(dt_tab2_wide, intervention)

for (col in names(dt_tab2_wide)[names(dt_tab2_wide) != "intervention"]) {
  dt_tab2_wide[, (col) := fifelse(
    is.na(get(col)), "\u2013",
    comma(round(get(col) / 1e3, 1), accuracy = 0.1)
  )]
}

kable(dt_tab2_wide,
      caption = "Table 2. Cumulative deaths averted (thousands, 2017\u20132050) by intervention and cause of death.",
      align   = c("l", rep("r", ncol(dt_tab2_wide) - 1))) |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = TRUE, font_size = 13) |>
  add_header_above(setNames(c(1L, ncol(dt_tab2_wide) - 1L),
                            c(" ", "Deaths averted, thousands (2017\u20132050)"))) |>
  row_spec(0, bold = TRUE) |>
  footnote(general = paste0(
    "\u2013 = intervention does not affect this cause. ",
    "TFA elimination affects IHD only. ",
    "Antihypertensive interventions affect all four CVD causes."
  ))

dt_excel_tab2 <- copy(dt_tab2_wide)
```

## Figure 2 --- Deaths Averted by Cause (Stacked Bar)

```{r figure2, message=FALSE, warning=FALSE, echo=FALSE, fig.width=9, fig.height=5}
dt_fig2 <- dt_merged[
  intervention %in% int_order & cause %in% names(cause_map),
  .(deaths_averted = sum(deaths_averted, na.rm = TRUE)),
  by = .(intervention, cause)
]
dt_fig2[, intervention := factor(intervention, levels = int_order)]
dt_fig2[, cause_lbl := cause_labels_plot[cause]]
dt_fig2[, cause_lbl := factor(cause_lbl, levels = unname(cause_labels_plot))]

g <- ggplot(dt_fig2, aes(x = deaths_averted / 1e3, y = intervention, fill = cause_lbl)) +
  geom_col(width = 0.65) +
  scale_fill_manual(values = setNames(cause_cols, cause_labels_plot),
                    name = "Cause of death") +
  scale_x_continuous(expand = c(0, 0), labels = comma) +
  guides(fill = guide_legend(nrow = 2, byrow = TRUE)) +
  labs(
    title = "Deaths Averted by Cause of Death and Intervention, 2017\u20132050",
    x     = "Deaths averted (thousands)",
    y     = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(legend.position    = "bottom",
        panel.grid.minor   = element_blank(),
        panel.grid.major.y = element_blank(),
        plot.title         = element_text(face = "bold"))
print(g)
ggsave(paste0(wd_outp, "assessment_fig2_deaths_cause.png"), plot = g,
       width = 9, height = 5, dpi = 300)
```

## Figure 3 --- Time Series: Annual Deaths Averted by Intervention

```{r figure3-annual, message=FALSE, warning=FALSE, echo=FALSE, fig.width=10, fig.height=5}
g <- ggplot(dt_annual[!intervention %in% c("HTN Control + TFA Elimination")],
            aes(x = year, y = deaths_averted / 1e3,
                color = intervention, fill = intervention)) +
  geom_area(position = "stack", alpha = 0.7, color = "grey20", linewidth = 0.15) +
  scale_color_manual(values = int_cols, name = NULL) +
  scale_fill_manual(values  = int_cols, name = NULL) +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Annual Program-Attributable Deaths Averted by Intervention, 2017\u20132050",
    x     = "Year",
    y     = "Deaths averted (thousands)"
  ) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "bottom", panel.grid.minor = element_blank(),
        plot.title = element_text(face = "bold"))
print(g)
ggsave(paste0(wd_outp, "assessment_fig3_annual_averted.png"), plot = g,
       width = 11, height = 5, dpi = 300)
```

## Figure 4 --- Cumulative Deaths Averted Over Time

```{r figure4-cumulative, message=FALSE, warning=FALSE, echo=FALSE, fig.width=10, fig.height=5}
g <- ggplot(dt_annual,
            aes(x = year, y = cum_deaths_averted / 1e3, color = intervention)) +
  geom_line(linewidth = 1.1) +
  scale_color_manual(values = int_cols, name = NULL) +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Cumulative Deaths Averted by Intervention, 2017\u20132050",
    x     = "Year",
    y     = "Cumulative deaths averted (thousands)"
  ) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "bottom", panel.grid.minor = element_blank(),
        plot.title = element_text(face = "bold"))
print(g)
ggsave(paste0(wd_outp, "assessment_fig4_cum_averted.png"), plot = g,
       width = 11, height = 5, dpi = 300)
```

---

# Outcomes per 1 Million

```{r per1m-compute, message=FALSE, warning=FALSE, echo=FALSE}
# ---- Reconstruct actual population denominators from raw inputs ----
# Replicating the logic from 05_build_baseline.R (not from model output)

# PROGRAM POPULATION (HTN denominator): actual enrolled per location-year
dt_program_pop <- tryCatch({
  # Step a: Load htn_control_scenarios, keep location, year, enrolled
  htn_csv <- fread(paste0(wd_data, "htn_control_scenarios_by_loc_year.csv"))
  htn_csv <- htn_csv[, .(location, year, enrolled)]

  # Step b: Compute age/sex population proportions from model output
  #         (data_bau, Counterfactual, cause == "ihd", year >= 2017)
  dt_pop_dist <- data_bau[
    intervention == "Counterfactual" & cause == "ihd" & year >= 2017,
    .(population = sum(pop, na.rm = TRUE)),
    by = .(location, year, age, sex)
  ]
  dt_pop_dist[, population_proportion := population / sum(population),
              by = .(location, year)]

  # Step c: Merge enrolled x population_proportion to get Nx_program
  dt_enr <- merge(htn_csv, dt_pop_dist,
                  by = c("location", "year"), all.y = TRUE)
  dt_enr[is.na(enrolled), enrolled := 0]
  dt_enr[, Nx_program := enrolled * population_proportion]

  # Step d: Sum by location-year to get total program_population
  dt_enr[, .(program_population = sum(Nx_program, na.rm = TRUE)),
         by = .(location, year)]
}, error = function(e) {
  message("htn_control_scenarios_by_loc_year.csv not found ",
          "\u2014 program_population unavailable.")
  NULL
})

# TOTAL COUNTRY POPULATION age 20+ per location-year
# (from UNWPP, same source as 05_build_baseline.R)
dt_total_pop <- tryCatch({
  pop_raw <- fread(paste0(wd_data, "PopulationsAge20_2050.csv"))
  if ("year_id" %in% names(pop_raw)) setnames(pop_raw, "year_id", "year")
  pop_raw[, .(total_population = sum(Nx, na.rm = TRUE)),
          by = .(location, year)]
}, error = function(e) {
  message("PopulationsAge20_2050.csv not found ",
          "\u2014 total_population unavailable.")
  NULL
})

# ---- Per-1M summary by intervention x cause ----
# Model outputs are already per 1M treated (upstream normalisation).
# HTN scenarios: per 1M treated patients
# TFA scenarios: per 1M population (approximately equal per-person rate)

dt_per1m_int_cause <- dt_merged[
  year >= 2017,
  .(deaths_per_1M        = round(sum(pgm_deaths,     na.rm = TRUE), 1),
    deaths_averted_per_1M = round(sum(deaths_averted, na.rm = TRUE), 1)),
  by = .(intervention, cause)
]
dt_per1m_int_cause[, cause_lbl := cause_map[cause]]
dt_per1m_int_cause[, denominator := fifelse(
  intervention == "TFA Elimination",
  "per 1M population", "per 1M treated"
)]

# Totals by intervention (for bar chart)
dt_per1m_bar <- dt_merged[
  year >= 2017,
  .(deaths_averted_per_1M = sum(deaths_averted, na.rm = TRUE)),
  by = intervention
]
dt_per1m_bar[, intervention := factor(intervention, levels = int_order)]

# ---- Rescale to actual absolute deaths averted ----
dt_rescale_detail <- dt_merged[
  year >= 2017,
  .(per_1M_averted = sum(deaths_averted, na.rm = TRUE)),
  by = .(intervention, location, year)
]

if (!is.null(dt_program_pop)) {
  dt_rescale_detail <- merge(dt_rescale_detail, dt_program_pop,
                             by = c("location", "year"), all.x = TRUE)
} else {
  dt_rescale_detail[, program_population := NA_real_]
}
if (!is.null(dt_total_pop)) {
  dt_rescale_detail <- merge(dt_rescale_detail, dt_total_pop,
                             by = c("location", "year"), all.x = TRUE)
} else {
  dt_rescale_detail[, total_population := NA_real_]
}

# HTN: multiply per-1M by (actual program_population / 1e6)
dt_rescale_detail[
  intervention %in% c("HTN Control (Program)",
                       "HTN Control + TFA Elimination"),
  abs_averted := per_1M_averted * (program_population / 1e6)
]
# TFA: multiply per-1M by (total_population / 1e6)
dt_rescale_detail[
  intervention == "TFA Elimination",
  abs_averted := per_1M_averted * (total_population / 1e6)
]

# Aggregate for comparison table
dt_rescale_agg <- dt_rescale_detail[,
  .(rescaled_abs = sum(abs_averted, na.rm = TRUE),
    model_per_1M = sum(per_1M_averted, na.rm = TRUE)),
  by = intervention
]
dt_rescale_agg[, model_per_1M_K  := round(model_per_1M  / 1e3, 1)]
dt_rescale_agg[, rescaled_abs_K  := round(rescaled_abs  / 1e3, 1)]

# Flag large discrepancies (> 10%) between rescaled and model per-1M
# (expected to differ because of population scaling; flag only unexpected mismatches)
dt_rescale_agg[, ratio := rescaled_abs / pmax(model_per_1M, 1)]

# ---- Additional per-1M scalars for narrative ----
# (lives_per_1M_htn/tfa/combined already computed in per1m-scalars chunk)

# 5-year window (2020-2024) average per country
lives_per_1M_htn_5yr <- round(
  dt_merged[intervention == "HTN Control (Program)" &
              year %in% 2020:2024,
            sum(deaths_averted, na.rm = TRUE)] / n_countries
)

# Rescaled absolute totals
total_bp_rescaled <- sum(
  dt_rescale_detail[intervention == "HTN Control (Program)",
                    abs_averted], na.rm = TRUE)
total_tfa_rescaled <- sum(
  dt_rescale_detail[intervention == "TFA Elimination",
                    abs_averted], na.rm = TRUE)
total_combined_rescaled <- sum(
  dt_rescale_detail[intervention == "HTN Control + TFA Elimination",
                    abs_averted], na.rm = TRUE)
```

## Per 1M Treated (Normalised Model Output)

```{r per1m-table, message=FALSE, warning=FALSE, echo=FALSE}
dt_per1m_display <- dt_per1m_int_cause[, .(
  Intervention              = intervention,
  Cause                     = cause_lbl,
  `Deaths per 1M`           = comma(deaths_per_1M,        accuracy = 0.1),
  `Deaths averted per 1M`   = comma(deaths_averted_per_1M, accuracy = 0.1),
  Denominator               = denominator
)]
setorder(dt_per1m_display, Intervention, Cause)

kable(dt_per1m_display,
      caption = paste0(
        "Deaths and deaths averted per 1 million, by intervention and ",
        "cause (", study_years, ")."),
      align = c("l", "l", "r", "r", "l")) |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = TRUE, font_size = 13) |>
  row_spec(0, bold = TRUE) |>
  footnote(general = paste0(
    "HTN scenarios: per 1M patients treated (normalised model output). ",
    "TFA scenarios: per 1M population age 20+. ",
    "Figures are cumulative sums across all ", n_countries,
    " program countries."))
```

```{r per1m-barchart, message=FALSE, warning=FALSE, echo=FALSE, fig.width=9, fig.height=5}
g <- ggplot(dt_per1m_bar,
            aes(x = deaths_averted_per_1M / 1e3,
                y = intervention,
                fill = as.character(intervention))) +
  geom_col(width = 0.65) +
  scale_fill_manual(values = int_cols, guide = "none") +
  scale_x_continuous(expand = c(0, 0), labels = comma) +
  labs(
    title = paste0("Deaths Averted per 1 Million by Intervention, ",
                   study_years),
    x     = "Deaths averted per 1M (thousands, cumulative across countries)",
    y     = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(plot.title         = element_text(face = "bold"),
        panel.grid.minor   = element_blank(),
        panel.grid.major.y = element_blank())
print(g)
```

## Rescaled to Actual Population

```{r per1m-rescaled, message=FALSE, warning=FALSE, echo=FALSE}
dt_rescale_display <- dt_rescale_agg[, .(
  Intervention                     = intervention,
  `Model per-1M (thousands)`       = comma(model_per_1M_K,  accuracy = 0.1),
  `Rescaled absolute (thousands)`  = comma(rescaled_abs_K,   accuracy = 0.1)
)]
dt_rescale_display[, Intervention := factor(Intervention, levels = int_order)]
setorder(dt_rescale_display, Intervention)

kable(dt_rescale_display,
      caption = paste0(
        "Rescaled absolute deaths averted vs. normalised per-1M model ",
        "output (", study_years, ")."),
      align = c("l", "r", "r")) |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE, font_size = 13) |>
  row_spec(0, bold = TRUE) |>
  footnote(general = paste0(
    "HTN rescaled by actual program enrollment; ",
    "TFA rescaled by total population age 20+. ",
    "Per-1M figures sum normalised model output across all countries."))
```

## Narrative Summary

The RTSL HTN program is projected to save approximately
`r fmt_k(lives_per_1M_htn)` thousand lives per million patients treated
(cumulative `r study_years`), equivalent to
`r fmt_k(total_bp_rescaled)` thousand deaths averted across all
`r n_countries` program countries. Over a 5-year window (2020--2024),
the average program country averts `r fmt_k(lives_per_1M_htn_5yr)` thousand
deaths per million treated.

TFA elimination policies are projected to avert
`r fmt_k(lives_per_1M_tfa)` thousand deaths per million population
(cumulative), rescaling to `r fmt_k(total_tfa_rescaled)` thousand actual
deaths averted across program countries.

Combined, `r fmt_k(lives_per_1M_combined)` thousand deaths per million
treated are averted, equivalent to `r fmt_k(total_combined_rescaled)` thousand
deaths averted in absolute terms.

---

# Outputs by Demographics

## By Sex and Age Group

```{r demo-age-sex, message=FALSE, warning=FALSE, echo=FALSE, fig.width=10, fig.height=5}
dt_age_sex <- dt_merged[
  intervention %in% int_order,
  .(deaths_averted = sum(deaths_averted, na.rm = TRUE)),
  by = .(intervention, age_group, sex)
]

order_lvls <- dt_age_sex[, .(total = sum(deaths_averted, na.rm = TRUE)), by = intervention
                         ][order(total), intervention]
dt_age_sex[, intervention := factor(intervention, levels = order_lvls)]

g <- ggplot(dt_age_sex,
            aes(x = deaths_averted / 1e3, y = intervention, fill = age_group)) +
  geom_col(width = 0.7) +
  facet_wrap(~sex, nrow = 1) +
  scale_x_continuous(expand = c(0, 0), labels = comma) +
  labs(
    title = "Deaths Averted by Age Group, Sex, and Intervention (2017\u20132050)",
    x     = "Deaths averted (thousands)",
    y     = NULL,
    fill  = "Age group"
  ) +
  theme_minimal(base_size = 13) +
  theme(legend.position  = "bottom",
        panel.border     = element_rect(color = "black", fill = NA, linewidth = 0.5),
        strip.text       = element_text(face = "bold"),
        panel.grid.minor = element_blank(),
        plot.title       = element_text(face = "bold"))
print(g)
ggsave(paste0(wd_outp, "assessment_fig5_demo_age.png"), plot = g,
       width = 10, height = 5, dpi = 300)
```

## Table 3 --- Deaths Averted by Age Group and Intervention

```{r table3-age, message=FALSE, warning=FALSE, echo=FALSE}
# Build age-group summary using program vs bau comparison
dt_tab3_base <- data_bau[year >= 2017 & intervention %in% int_order,
                          .(baseline_deaths = sum(dead, na.rm = TRUE)),
                          by = .(age_group, intervention)]
dt_tab3_pgm  <- data_program[year >= 2017 & intervention %in% int_order,
                              .(pgm_deaths = sum(dead, na.rm = TRUE)),
                              by = .(age_group, intervention)]

dt_tab3 <- merge(dt_tab3_pgm, dt_tab3_base,
                 by = c("age_group", "intervention"), all = TRUE)
dt_tab3[, deaths_averted    := baseline_deaths - pgm_deaths]
dt_tab3[, deaths_averted_yr := deaths_averted / n_years]
dt_tab3[, pct_averted       := deaths_averted / baseline_deaths]

dt_tab3[, intervention := factor(intervention, levels = int_order)]
setorder(dt_tab3, age_group, intervention)

dt_tab3_fmt <- dt_tab3[, .(
  Age_group          = age_group,
  Intervention       = intervention,
  Counterfactual     = comma(baseline_deaths,   accuracy = 1),
  Program            = comma(pgm_deaths,        accuracy = 1),
  Deaths_averted     = comma(deaths_averted,    accuracy = 1),
  Deaths_averted_yr  = comma(deaths_averted_yr, accuracy = 1),
  Pct_averted        = percent(pct_averted, accuracy = 0.1)
)]
setnames(dt_tab3_fmt,
         old = c("Age_group", "Counterfactual", "Program",
                 "Deaths_averted", "Deaths_averted_yr", "Pct_averted"),
         new = c("Age group", "Counterfactual deaths", "Program deaths",
                 "Deaths averted", "Deaths averted/year", "% averted"))

kable(dt_tab3_fmt,
      caption = "Table 3. Deaths averted (2017\u20132050) by age group and intervention.",
      align   = c("l","l","r","r","r","r","r")) |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = TRUE, font_size = 12) |>
  row_spec(0, bold = TRUE)

dt_excel_tab3 <- copy(dt_tab3_fmt)
```

---

# Outputs by Geography

## Table 4 --- Deaths Averted by Country (HTN + TFA Combined)

```{r table4-country, message=FALSE, warning=FALSE, echo=FALSE}
dt_tab4_base <- data_bau[year >= 2017 &
                           intervention == "HTN Control + TFA Elimination",
                          .(baseline_deaths = sum(dead, na.rm = TRUE)),
                          by = .(region, location)]
dt_tab4_pgm  <- data_program[year >= 2017 &
                               intervention == "HTN Control + TFA Elimination",
                              .(pgm_deaths = sum(dead, na.rm = TRUE)),
                              by = .(region, location)]

dt_tab4 <- merge(dt_tab4_pgm, dt_tab4_base, by = c("region", "location"), all = TRUE)
dt_tab4[, deaths_averted    := baseline_deaths - pgm_deaths]
dt_tab4[, deaths_averted_yr := deaths_averted / n_years]
dt_tab4[, pct_averted       := deaths_averted / baseline_deaths]
setorder(dt_tab4, region, -deaths_averted)

dt_tab4_fmt <- dt_tab4[, .(
  Region             = region,
  Country            = location,
  Counterfactual     = comma(baseline_deaths,    accuracy = 1),
  Program            = comma(pgm_deaths,         accuracy = 1),
  Deaths_averted     = comma(deaths_averted,     accuracy = 1),
  Deaths_averted_yr  = comma(deaths_averted_yr,  accuracy = 1),
  Pct_averted        = percent(pct_averted, accuracy = 0.1)
)]
setnames(dt_tab4_fmt,
         old = c("Counterfactual", "Program",
                 "Deaths_averted", "Deaths_averted_yr", "Pct_averted"),
         new = c("Counterfactual deaths", "Program deaths",
                 "Deaths averted", "Deaths averted/year", "% averted"))

datatable(
  dt_tab4_fmt,
  caption  = htmltools::tags$caption(
    style = "caption-side: top; text-align: left;",
    "Table 4. Country-level program-attributable deaths averted (HTN + TFA combined), 2017\u20132050."
  ),
  filter   = "top",
  rownames = FALSE,
  options  = list(pageLength = 15, scrollX = TRUE, dom = "lftipr")
)

dt_excel_tab4 <- copy(dt_tab4_fmt)
```

## Figure 5 --- Top Countries by Deaths Averted

```{r figure5-top-countries, message=FALSE, warning=FALSE, echo=FALSE, fig.width=10, fig.height=7}
n_show <- min(n_countries, 20)
dt_top <- dt_tab4[order(-deaths_averted)][1:n_show]
dt_top[, location := factor(location, levels = rev(location))]

g <- ggplot(dt_top, aes(x = deaths_averted / 1e3, y = location, fill = region)) +
  geom_col(width = 0.7) +
  scale_x_continuous(expand = c(0, 0), labels = comma) +
  labs(
    title = paste0("Top ", n_show, " Countries by Program-Attributable Deaths Averted\n",
                   "(HTN + TFA, 2017\u20132050)"),
    x     = "Deaths averted (thousands)",
    y     = NULL,
    fill  = "Region"
  ) +
  theme_minimal(base_size = 13) +
  theme(legend.position    = "right",
        panel.grid.minor   = element_blank(),
        panel.grid.major.y = element_blank(),
        plot.title         = element_text(face = "bold"))
print(g)
ggsave(paste0(wd_outp, "assessment_fig5_top_countries.png"), plot = g,
       width = 10, height = 7, dpi = 300)
```

## Figure 6 --- World Map: Deaths Averted per 100,000 Population

```{r figure6-map, message=FALSE, warning=FALSE, echo=FALSE, fig.width=11, fig.height=5}
wpp <- tryCatch(
  as.data.table(readRDS(paste0(wd_data, "PopulationsSingleAge0050.rds"))),
  error = function(e) {
    tryCatch(
      as.data.table(readRDS(paste0(wd_data, "UN2024/PopulationsSingleAge0050.rds"))),
      error = function(e2) NULL
    )
  }
)

if (!is.null(wpp)) {
  pop_2019 <- wpp[year_id == 2019 & age >= 20,
                  .(pop = sum(Nx, na.rm = TRUE)), by = location]
  dt_map <- merge(dt_tab4, pop_2019, by = "location", all.x = TRUE)
  dt_map[, da_per100k := deaths_averted / pop * 1e5]

  if (nrow(countries_dt) > 0) {
    world    <- ne_countries(scale = "medium", returnclass = "sf")
    world_dt <- merge(world,
                      merge(dt_map, countries_dt[, .(location, iso3)],
                            by = "location", all.x = TRUE),
                      by.x = "adm0_a3", by.y = "iso3", all.x = TRUE)

    g <- ggplot(world_dt) +
      geom_sf(aes(fill = da_per100k), color = "grey80", linewidth = 0.1) +
      scale_fill_viridis_c(name = "per 100,000", na.value = "grey90",
                           option = "C", direction = -1) +
      labs(
        title    = "Program-Attributable Deaths Averted per 100,000 Population, 2017\u20132050",
        subtitle = "HTN control + TFA elimination combined",
        caption  = "Source: Model output; denominator = UNWPP 2024 age 20+ population"
      ) +
      theme_void(base_size = 12) +
      theme(legend.position = "right",
            plot.title    = element_text(face = "bold"),
            plot.subtitle = element_text(color = "grey40"))
    print(g)
    ggsave(paste0(wd_outp, "assessment_fig6_map.png"), plot = g,
           width = 11, height = 5, dpi = 300)
  }
} else {
  message("Population file not found - skipping map.")
}
```

## Table 5 --- Deaths Averted by Region

```{r table5-region, message=FALSE, warning=FALSE, echo=FALSE}
dt_tab5_base <- data_bau[year >= 2017 &
                           intervention == "HTN Control + TFA Elimination",
                          .(baseline_deaths = sum(dead, na.rm = TRUE)), by = region]
dt_tab5_pgm  <- data_program[year >= 2017 &
                               intervention == "HTN Control + TFA Elimination",
                              .(pgm_deaths = sum(dead, na.rm = TRUE)), by = region]

dt_tab5 <- merge(dt_tab5_pgm, dt_tab5_base, by = "region", all = TRUE)
dt_tab5[, deaths_averted    := baseline_deaths - pgm_deaths]
dt_tab5[, deaths_averted_yr := deaths_averted / n_years]
dt_tab5[, pct_averted       := deaths_averted / baseline_deaths]
setorder(dt_tab5, -deaths_averted)

dt_tab5_fmt <- dt_tab5[, .(
  Region             = region,
  Counterfactual     = comma(baseline_deaths,    accuracy = 1),
  Program            = comma(pgm_deaths,         accuracy = 1),
  Deaths_averted     = comma(deaths_averted,     accuracy = 1),
  Deaths_averted_yr  = comma(deaths_averted_yr,  accuracy = 1),
  Pct_averted        = percent(pct_averted, accuracy = 0.1)
)]
setnames(dt_tab5_fmt,
         old = c("Counterfactual", "Program",
                 "Deaths_averted", "Deaths_averted_yr", "Pct_averted"),
         new = c("Counterfactual deaths", "Program deaths",
                 "Deaths averted", "Deaths averted/year", "% averted"))

kable(dt_tab5_fmt,
      caption = "Table 5. Deaths averted by region (HTN + TFA combined), 2017\u20132050.",
      align   = c("l","r","r","r","r","r")) |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE, font_size = 13) |>
  row_spec(0, bold = TRUE)

dt_excel_tab5 <- copy(dt_tab5_fmt)
```

---

# ASMR Under Program vs Counterfactual

```{r asmr-interventions, message=FALSE, warning=FALSE, echo=FALSE, fig.width=12, fig.height=7}
# Compute ASMR for program and counterfactual runs (combined intervention = HTN+TFA)
dt_asmr_pgm <- data_program[
  intervention == "HTN Control + TFA Elimination" & year >= 2017,
  .(dead = sum(dead, na.rm = TRUE),
    well = sum(well, na.rm = TRUE),
    sick = sum(sick, na.rm = TRUE)),
  by = .(age, year)]
dt_asmr_pgm[, Nx := well + sick]
dt_asmr_pgm[, scenario := "Program (HTN + TFA)"]

dt_asmr_bau <- data_bau[
  intervention == "HTN Control + TFA Elimination" & year >= 2017,
  .(dead = sum(dead, na.rm = TRUE),
    well = sum(well, na.rm = TRUE),
    sick = sum(sick, na.rm = TRUE)),
  by = .(age, year)]
dt_asmr_bau[, Nx := well + sick]
dt_asmr_bau[, scenario := "Counterfactual"]

dt_asmr_both <- rbind(dt_asmr_pgm, dt_asmr_bau)
asmr_int <- calculate_asmr(dt_asmr_both, deaths_col = "dead",
                            population_col = "Nx", disagg_var = "scenario")

g <- ggplot(asmr_int,
            aes(x = year, y = ASMR, color = scenario, linetype = year > 2024)) +
  geom_line(linewidth = 1.1) +
  scale_color_manual(values = c("Counterfactual"       = "#999999",
                                "Program (HTN + TFA)"  = "#6a3d9a"),
                     name = NULL) +
  scale_linetype_manual(values = c(`FALSE` = "solid", `TRUE` = "dashed"), guide = "none") +
  labs(
    title    = "Age-Standardised CVD Mortality Rate: Program vs Counterfactual",
    subtitle = paste0("All causes combined; solid = historical, dashed = projected; ",
                      n_countries, " RTSL countries"),
    x        = "Year",
    y        = "ASMR (per 100,000 population)",
    caption  = "WHO World Standard Population (2000\u20132025), rescaled to ages 20\u201395+"
  ) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "bottom", panel.grid.minor = element_blank(),
        plot.title = element_text(face = "bold"))
print(g)
ggsave(paste0(wd_outp, "assessment_asmr_interventions.png"), plot = g,
       width = 12, height = 6, dpi = 300)
```

---

# Premature Mortality (Deaths < 70 years)

```{r premature-mortality, message=FALSE, warning=FALSE, echo=FALSE}
dt_u70 <- dt_merged[intervention %in% int_order & age_group == "<70",
                     .(deaths_averted = sum(deaths_averted, na.rm = TRUE)),
                     by = .(intervention, sex)]

dt_u70_total <- dt_merged[intervention == "HTN Control + TFA Elimination" &
                            age_group == "<70",
                          .(deaths_averted_u70 = sum(deaths_averted, na.rm = TRUE)),
                          by = location]

dt_u70_all <- dt_merged[intervention == "HTN Control + TFA Elimination",
                         .(deaths_averted_total = sum(deaths_averted, na.rm = TRUE)),
                         by = location]

dt_u70_loc <- merge(dt_u70_total, dt_u70_all, by = "location")
dt_u70_loc[, prop_u70 := fifelse(deaths_averted_total == 0, NA_real_,
                                  deaths_averted_u70 / deaths_averted_total)]
setorder(dt_u70_loc, -deaths_averted_u70)

kable(dt_u70[, .(Intervention = intervention, Sex = sex,
                  `Deaths averted <70` = comma(deaths_averted, accuracy = 1))],
      caption = "Deaths averted among people under 70, by intervention and sex (2017\u20132050).",
      align = "llr") |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE, font_size = 13) |>
  row_spec(0, bold = TRUE)
```

```{r premature-country-table, message=FALSE, warning=FALSE, echo=FALSE}
dt_u70_fmt <- dt_u70_loc[, .(
  Country = location,
  `Deaths averted <70` = comma(deaths_averted_u70, accuracy = 1),
  `Share of total averted` = percent(prop_u70, accuracy = 0.1)
)]

datatable(
  dt_u70_fmt,
  caption  = htmltools::tags$caption(
    style = "caption-side: top; text-align: left;",
    "Deaths averted under age 70 by country (HTN + TFA combined, 2017\u20132050)."
  ),
  filter   = "top",
  rownames = FALSE,
  options  = list(pageLength = 15, scrollX = TRUE, dom = "lftipr")
)
```

---

# Excel Export

```{r excel-export, message=FALSE, warning=FALSE, echo=FALSE}
wb <- createWorkbook()

add_sheet <- function(wb, sheet_name, dat, title_str) {
  addWorksheet(wb, sheet_name)
  writeData(wb, sheet_name, title_str, startRow = 1, startCol = 1)
  addStyle(wb, sheet_name,
           style = createStyle(fontSize = 12, textDecoration = "bold"),
           rows = 1, cols = 1)
  writeData(wb, sheet_name, dat, startRow = 3, startCol = 1,
            headerStyle = createStyle(textDecoration = "bold",
                                      fgFill = "#2E75B6", fontColour = "white",
                                      border = "Bottom"))
  setColWidths(wb, sheet_name, cols = seq_len(ncol(dat)), widths = "auto")
}

add_sheet(wb, "T1_Deaths_by_Intervention", as.data.frame(dt_excel_tab1),
          "Table 1. Program-attributable deaths averted (2017-2050) by intervention.")
add_sheet(wb, "T2_Deaths_by_Cause", as.data.frame(dt_excel_tab2),
          "Table 2. Deaths averted (thousands) by intervention and cause of death.")
add_sheet(wb, "T3_Deaths_by_Age", as.data.frame(dt_excel_tab3),
          "Table 3. Deaths averted by age group and intervention.")
add_sheet(wb, "T4_Deaths_by_Country", as.data.frame(dt_excel_tab4),
          "Table 4. Country-level deaths averted (HTN + TFA, 2017-2050).")

if (exists("dt_excel_tab5") && nrow(dt_excel_tab5) > 0)
  add_sheet(wb, "T5_Deaths_by_Region", as.data.frame(dt_excel_tab5),
            "Table 5. Deaths averted by region (HTN + TFA, 2017-2050).")

add_sheet(wb, "Countries", as.data.frame(countries_by_region),
          "Countries included in the assessment, by region.")

# Premature mortality
if (exists("dt_u70_loc") && nrow(dt_u70_loc) > 0)
  add_sheet(wb, "U70_by_Country", as.data.frame(dt_u70_loc),
            "Deaths averted under age 70 by country (HTN + TFA, 2017-2050).")

# ---- Per1M_Summary sheet ----
# One row per location x year x scenario, all scenarios stacked
dt_cf_ly <- data_bau[
  intervention == "Counterfactual" & year >= 2017,
  .(deaths = sum(dead, na.rm = TRUE)),
  by = .(location, year)
]
dt_cf_ly[, scenario := "Counterfactual"]

dt_int_ly <- dt_merged[
  year >= 2017,
  .(deaths         = sum(pgm_deaths,     na.rm = TRUE),
    deaths_averted = sum(deaths_averted, na.rm = TRUE)),
  by = .(location, year, intervention)
]
setnames(dt_int_ly, "intervention", "scenario")

# Counterfactual deaths reference (for pct_averted)
dt_cf_ref_ly <- dt_cf_ly[, .(location, year, cf_deaths_ref = deaths)]

dt_excel_per1m <- rbind(
  dt_cf_ly[, .(location, year, scenario, deaths, deaths_averted = 0)],
  dt_int_ly[, .(location, year, scenario, deaths, deaths_averted)],
  fill = TRUE
)
dt_excel_per1m <- merge(dt_excel_per1m, dt_cf_ref_ly,
                        by = c("location", "year"), all.x = TRUE)

# Add population denominators
if (!is.null(dt_total_pop)) {
  dt_excel_per1m <- merge(dt_excel_per1m, dt_total_pop,
                          by = c("location", "year"), all.x = TRUE)
} else {
  dt_excel_per1m[, total_population := NA_real_]
}

if (!is.null(dt_program_pop)) {
  dt_excel_per1m <- merge(dt_excel_per1m, dt_program_pop,
                          by = c("location", "year"), all.x = TRUE)
} else {
  dt_excel_per1m[, program_population := NA_real_]
}

# For TFA and Counterfactual: program_population = total_population
dt_excel_per1m[
  scenario %in% c("TFA Elimination", "Counterfactual"),
  program_population := total_population
]

# Rename total_population -> population
setnames(dt_excel_per1m, "total_population", "population")

# Compute per-1M columns (model output is already per 1M; values are direct)
dt_excel_per1m[, deaths_per_1M         := round(deaths, 1)]
dt_excel_per1m[, deaths_averted_per_1M := round(deaths_averted, 1)]
dt_excel_per1m[scenario == "Counterfactual", deaths_averted_per_1M := 0]

# Absolute deaths averted (rescaled)
dt_excel_per1m[
  scenario %in% c("HTN Control (Program)",
                   "HTN Control + TFA Elimination"),
  n_deaths_averted := deaths_averted * (program_population / 1e6)
]
dt_excel_per1m[
  scenario == "TFA Elimination",
  n_deaths_averted := deaths_averted * (population / 1e6)
]
dt_excel_per1m[scenario == "Counterfactual", n_deaths_averted := 0]

# pct_averted
dt_excel_per1m[, pct_averted := round(
  fifelse(cf_deaths_ref > 0 & scenario != "Counterfactual",
          deaths_averted / cf_deaths_ref * 100, 0), 2)]
dt_excel_per1m[scenario == "Counterfactual", pct_averted := 0]

# Select final columns and sort
dt_excel_per1m_final <- dt_excel_per1m[, .(
  location, year, scenario, population, program_population,
  deaths_per_1M, deaths_averted_per_1M, n_deaths_averted, pct_averted
)]
setorder(dt_excel_per1m_final, location, scenario, year)

add_sheet(wb, "Per1M_Summary", as.data.frame(dt_excel_per1m_final),
          "Per 1M Summary: deaths and deaths averted by location, year, and scenario.")

out_xlsx <- paste0(wd_scen, "RTSL_Assessment_Results.xlsx")
saveWorkbook(wb, out_xlsx, overwrite = TRUE)
message("Tables exported to: ", out_xlsx)
```

---

# Slide Deck Outputs

```{r slides-setup, message=FALSE, warning=FALSE, echo=FALSE}
slides_dir <- paste0(wd_outp, "slides/")
if (!dir.exists(slides_dir)) dir.create(slides_dir, recursive = TRUE)
message("Slide artefacts will be written to: ", slides_dir)
```

```{r slides-scalars, message=FALSE, warning=FALSE, echo=FALSE}
bau_total_deaths <- data_bau[intervention == "HTN Control + TFA Elimination" &
                               year >= 2017,
                              sum(dead, na.rm = TRUE)]
sl_scalars <- list(
  n_countries      = n_countries,
  study_years      = study_years,
  year_min         = year_min,
  year_max         = year_max,
  n_years          = n_years,
  total_combined   = total_combined,
  total_bp         = total_bp,
  total_tfa        = total_tfa,
  bau_total_deaths = bau_total_deaths,
  lives_per_1M_htn      = lives_per_1M_htn,
  lives_per_1M_tfa      = lives_per_1M_tfa,
  lives_per_1M_combined = lives_per_1M_combined
)
saveRDS(sl_scalars, paste0(slides_dir, "sl_scalars.rds"))
message("SL-01 saved: sl_scalars.rds")
```

```{r slides-tab-summary, message=FALSE, warning=FALSE, echo=FALSE}
sl_tab_summary <- copy(dt_cumul)
sl_tab_summary[, pct    := deaths_averted / bau_total_deaths * 100]
sl_tab_summary[, cum_K  := paste0(fmt_k(deaths_averted), " K")]
sl_tab_summary[, avg_yr := paste0(fmt_k(deaths_averted / n_years), " K")]
sl_tab_summary[, pct_s  := paste0(round(pct, 1), "%")]
sl_tab_summary[, intervention := factor(intervention, levels = int_order)]
setorder(sl_tab_summary, intervention)
saveRDS(sl_tab_summary, paste0(slides_dir, "sl_tab_summary.rds"))
message("SL-02 saved: sl_tab_summary.rds")
```

```{r slides-tab-region, message=FALSE, warning=FALSE, echo=FALSE}
sl_tab_region <- merge(
  data_program[intervention == "HTN Control + TFA Elimination" & year >= 2017,
               .(pgm = sum(dead, na.rm = TRUE)), by = region],
  data_bau[intervention == "HTN Control + TFA Elimination" & year >= 2017,
           .(bau = sum(dead, na.rm = TRUE)), by = region],
  by = "region"
)
sl_tab_region[, averted := bau - pgm]
sl_tab_region[, pct     := averted / bau * 100]
setorder(sl_tab_region, -averted)
saveRDS(sl_tab_region, paste0(slides_dir, "sl_tab_region.rds"))
message("SL-03 saved: sl_tab_region.rds")
```

```{r slides-tab-top15, message=FALSE, warning=FALSE, echo=FALSE}
sl_tab_top15 <- merge(
  data_program[intervention == "HTN Control + TFA Elimination" & year >= 2017,
               .(pgm_dead = sum(dead, na.rm = TRUE),
                 region   = region[1]), by = location],
  data_bau[intervention == "HTN Control + TFA Elimination" & year >= 2017,
           .(bau = sum(dead, na.rm = TRUE)), by = location],
  by = "location"
)
sl_tab_top15[, averted := bau - pgm_dead]
setorder(sl_tab_top15, -averted)
sl_tab_top15 <- sl_tab_top15[1:min(15, .N)]
saveRDS(sl_tab_top15, paste0(slides_dir, "sl_tab_top15.rds"))
message("SL-04 saved: sl_tab_top15.rds")
```

```{r slides-dt-cause, message=FALSE, warning=FALSE, echo=FALSE}
sl_dt_cause <- dt_merged[intervention %in% int_order & cause %in% names(cause_map),
                          .(averted = sum(deaths_averted, na.rm = TRUE)),
                          by = .(cause, intervention)]
sl_dt_cause[, cause_lbl := fcase(
  cause == "ihd",     "IHD",
  cause == "istroke", "Ischemic stroke",
  cause == "hstroke", "Hemorrhagic stroke",
  cause == "hhd",     "Hypertensive HD",
  default = cause
)]
sl_dt_cause[, intervention := factor(intervention, levels = int_order)]
saveRDS(sl_dt_cause, paste0(slides_dir, "sl_dt_cause.rds"))
message("SL-05 saved: sl_dt_cause.rds")
```

```{r slides-dt-annual, message=FALSE, warning=FALSE, echo=FALSE}
saveRDS(copy(dt_annual)[intervention %in% int_order],
        paste0(slides_dir, "sl_dt_annual.rds"))
message("SL-06 saved: sl_dt_annual.rds")
```

```{r slides-fig-deaths-averted, message=FALSE, warning=FALSE, echo=FALSE}
g_ann_sl <- ggplot(dt_annual[!intervention %in% c("HTN Control + TFA Elimination")],
                    aes(x = year, y = deaths_averted / 1e3, fill = intervention)) +
  geom_area(alpha = 0.85, position = "stack") +
  scale_fill_manual(values = int_cols[int_order], name = NULL) +
  scale_y_continuous(labels = comma) +
  labs(x = "Year", y = "Deaths averted (thousands)") +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom",
        legend.text = element_text(size = 7),
        panel.grid.minor = element_blank()) +
  guides(fill = guide_legend(nrow = 2))
ggsave(paste0(slides_dir, "sl_fig_deaths_averted.png"), plot = g_ann_sl,
       width = 8, height = 4, dpi = 200)
message("SL-07 saved: sl_fig_deaths_averted.png")
```

```{r slides-fig-top15, message=FALSE, warning=FALSE, echo=FALSE}
sl_t15_plot <- copy(sl_tab_top15)
sl_t15_plot[, location := factor(location, levels = rev(location))]
g_top15_sl <- ggplot(sl_t15_plot,
                      aes(x = averted / 1e3, y = location, fill = region)) +
  geom_col(width = 0.7) +
  scale_x_continuous(labels = comma, expand = c(0, 0)) +
  labs(x = "Deaths averted (thousands)", y = NULL, fill = "Region") +
  theme_minimal(base_size = 10) +
  theme(legend.position = "right",
        panel.grid.minor = element_blank(),
        panel.grid.major.y = element_blank(),
        axis.text.y = element_text(size = 8))
ggsave(paste0(slides_dir, "sl_fig_top15_countries.png"), plot = g_top15_sl,
       width = 8, height = 5, dpi = 200)
message("SL-08 saved: sl_fig_top15_countries.png")
```

```{r slides-fig-cumulative, message=FALSE, warning=FALSE, echo=FALSE, fig.width=10, fig.height=5}
g <- ggplot(dt_annual,
            aes(x = year, y = cum_deaths_averted / 1e3, color = intervention)) +
  geom_line(linewidth = 1.1) +
  scale_color_manual(values = int_cols, name = NULL) +
  scale_y_continuous(labels = comma) +
  labs(x = "Year", y = "Cumulative deaths averted (thousands)") +
  theme_minimal(base_size = 13) +
  theme(legend.position = "bottom", panel.grid.minor = element_blank(),
        plot.title = element_text(face = "bold"))
print(g)
ggsave(paste0(slides_dir, "sl_fig_cum_averted.png"), plot = g,
       width = 11, height = 5, dpi = 300)
message("SL-09 saved: sl_fig_cum_averted.png")
```

```{r slides-bangladesh, message=FALSE, warning=FALSE, echo=FALSE}
# ---- Bangladesh country case artefact for presentation ---------------------
bgd_rds <- paste0(wd_outp, "out_model/model_output_Bangladesh_program_impact.rds")
if (file.exists(bgd_rds)) {
  dt_bgd <- setDT(readRDS(bgd_rds))

  # Program-attributable deaths by scenario
  dt_bgd_pgm <- dt_bgd[rate_scenario == "program" & year >= 2017,
                         .(pgm_deaths = sum(dead, na.rm = TRUE)),
                         by = .(scenario, year)]
  dt_bgd_cf  <- dt_bgd[rate_scenario == "bau_upper" &
                         scenario == "baseline" & year >= 2017,
                         .(cf_deaths = sum(dead, na.rm = TRUE)),
                         by = year]

  dt_bgd_merge <- merge(dt_bgd_pgm, dt_bgd_cf, by = "year")
  dt_bgd_merge[, averted := cf_deaths - pgm_deaths]

  # Summary table: total averted by scenario
  sl_bgd_summary <- dt_bgd_merge[, .(total_averted = sum(averted)), by = scenario]

  # Annual time-series for Bangladesh
  sl_bgd_annual <- dt_bgd_merge[, .(year, scenario, averted)]
  sl_bgd_annual[, cum_averted := cumsum(averted), by = scenario]

  # By cause
  dt_bgd_cause_pgm <- dt_bgd[rate_scenario == "program" & year >= 2017,
                               .(pgm_deaths = sum(dead, na.rm = TRUE)),
                               by = .(scenario, cause)]
  dt_bgd_cause_cf  <- dt_bgd[rate_scenario == "bau_upper" &
                               scenario == "baseline" & year >= 2017,
                               .(cf_deaths = sum(dead, na.rm = TRUE)),
                               by = cause]
  sl_bgd_cause <- merge(dt_bgd_cause_pgm, dt_bgd_cause_cf, by = "cause")
  sl_bgd_cause[, averted := cf_deaths - pgm_deaths]

  sl_bangladesh <- list(
    summary = sl_bgd_summary,
    annual  = sl_bgd_annual,
    cause   = sl_bgd_cause
  )
  saveRDS(sl_bangladesh, paste0(slides_dir, "sl_bangladesh.rds"))
  message("SL-10 saved: sl_bangladesh.rds")
} else {
  message("Bangladesh model output not found — skipping sl_bangladesh.rds")
}
```

```{r slides-tab-scenario, message=FALSE, warning=FALSE, echo=FALSE}
# ---- Scenario comparison table for presentation ----------------------------
sl_tab_scenario <- data.table(
  Intervention = c("HTN Control (Program)", "TFA Elimination",
                    "HTN Control + TFA Elimination"),
  Mechanism = c(
    "Antihypertensive therapy using observed RTSL program control rates",
    "Elimination of industrial TFA via enacted best-practice policies",
    "Both interventions combined (multiplicative effect ratios)"
  ),
  Key_Param = c(
    "Observed control rates (2017-2024); maintained at 2024 level to 2050",
    "Target: 0% energy from TFA; 2-year lag from policy passage to effect",
    "epsilon_IR_total = epsilon_IR_BP x epsilon_IR_TFA"
  ),
  Deaths_averted_K = c(
    round(total_bp / 1e3, 1),
    round(total_tfa / 1e3, 1),
    round(total_combined / 1e3, 1)
  )
)
saveRDS(sl_tab_scenario, paste0(slides_dir, "sl_tab_scenario.rds"))
message("SL-11 saved: sl_tab_scenario.rds")
```

```{r slides-manifest, message=FALSE, warning=FALSE, echo=FALSE}
sl_manifest <- data.table(
  file = c(
    "sl_scalars.rds", "sl_tab_summary.rds", "sl_tab_region.rds",
    "sl_tab_top15.rds", "sl_dt_cause.rds", "sl_dt_annual.rds",
    "sl_fig_deaths_averted.png", "sl_fig_top15_countries.png",
    "sl_fig_cum_averted.png",
    "sl_bangladesh.rds", "sl_tab_scenario.rds"
  ),
  type = c(rep("rds", 6), rep("png", 3), rep("rds", 2)),
  consumed_by = c(
    "inline scalars", "tab-summary kable", "tab-region kable",
    "fig-top-countries ggplot", "fig-by-cause ggplot", "fig-deaths-averted ggplot",
    "fig-deaths-averted include_graphics",
    "fig-top-countries include_graphics",
    "fig-cum-averted include_graphics",
    "bangladesh-results kable (presentation.Rmd)",
    "scenario-table kable (presentation.Rmd)"
  )
)
saveRDS(sl_manifest, paste0(slides_dir, "sl_manifest.rds"))

kable(sl_manifest,
      caption = "Slide deck artefact manifest. All files written to output/slides/.",
      col.names = c("File", "Type", "Consumed by (slides chunk)")) |>
  kable_styling(bootstrap_options = c("striped", "condensed"),
                full_width = TRUE, font_size = 11) |>
  row_spec(0, bold = TRUE)

message(sprintf("SLIDE DECK OUTPUTS complete. %d artefacts written to %s",
                nrow(sl_manifest), slides_dir))
```

---

*Report generated on `r Sys.time()` using R `r R.version$major`.`r R.version$minor`.*
