---
title: "RTSL Cardiovascular Health Program Impact Assessment"
subtitle: "Projected Health Outcomes 2017--2050"
author: "University of Washington -- Resolve to Save Lives"
institute: "\\color{red}Work in Progress"
date: "`r format(Sys.Date(), '%B %Y')`"
output:
  beamer_presentation:
    latex_engine: xelatex
    keep_tex: false
    slide_level: 2
    includes:
      in_header: beamer_preamble.tex
classoption: "aspectratio=169"
bibliography: references.bib
---

<!-- ======================================================================
     SETUP — LOAD ARTEFACTS ONLY (zero recomputation)
     All objects created by report.Rmd and saved to output/ and output/slides/
     ====================================================================== -->

```{r setup, include=FALSE}
library(data.table)
library(ggplot2)
library(knitr)
library(kableExtra)
library(scales)

# ---- Paths ------------------------------------------------------------------
#wd       <- normalizePath(file.path(dirname(getwd())), winslash = "/")
# Fallback: hardcode if knitting from docs/
#if (!dir.exists(file.path(wd, "output"))) {
 wd <- "C:/Users/wrgar/OneDrive - UW/02Work/ResolveToSaveLives/Assessment/uw-rtsl-assessment/"
#}
wd_data  <- paste0(wd, "data/processed/")
wd_outp  <- paste0(wd, "output/")
slides_dir <- paste0(wd_outp, "slides/")

# ---- Load pre-built slide artefacts -----------------------------------------
# These are generated by report.Rmd — knit report.Rmd first if missing.
if (file.exists(paste0(slides_dir, "sl_scalars.rds"))) {
  sl <- readRDS(paste0(slides_dir, "sl_scalars.rds"))
} else {
  sl <- list(n_countries = "N", study_years = "20XX--20XX")
}

n_countries           <- if (!is.null(sl$n_countries))      sl$n_countries      else "N"
study_years           <- if (!is.null(sl$study_years))      sl$study_years      else "20XX--20XX"
total_bp              <- if (!is.null(sl$total_bp))         sl$total_bp         else NA
total_tfa             <- if (!is.null(sl$total_tfa))        sl$total_tfa        else NA
total_combined        <- if (!is.null(sl$total_combined))   sl$total_combined   else NA
lives_per_1M_htn      <- if (!is.null(sl$lives_per_1M_htn))      sl$lives_per_1M_htn      else NA
lives_per_1M_tfa      <- if (!is.null(sl$lives_per_1M_tfa))      sl$lives_per_1M_tfa      else NA
lives_per_1M_combined <- if (!is.null(sl$lives_per_1M_combined)) sl$lives_per_1M_combined else NA

# ---- SL-02 to SL-08: Data tables -------------------------------------------
load_sl <- function(name) {
  f <- paste0(slides_dir, name)
  if (file.exists(f)) readRDS(f) else NULL
}

sl_tab_summary  <- load_sl("sl_tab_summary.rds")
sl_tab_region   <- load_sl("sl_tab_region.rds")
sl_tab_top15    <- load_sl("sl_tab_top15.rds")
sl_dt_cause     <- load_sl("sl_dt_cause.rds")
sl_dt_annual    <- load_sl("sl_dt_annual.rds")
sl_bangladesh   <- load_sl("sl_bangladesh.rds")
sl_tab_scenario <- load_sl("sl_tab_scenario.rds")

# ---- New slide artefacts (Data Sources & Key Variables section) ----
fig_htn_control_rates      <- load_sl("sl_fig_htn_control_rates.rds")
tbl_tfa_countries           <- load_sl("sl_tbl_tfa_countries.rds")
tbl_htn_program_vs_country <- load_sl("sl_tbl_htn_program_vs_country.rds")
tbl_countries_by_region    <- load_sl("sl_tbl_countries_by_region.rds")

# ---- New slide artefacts (Results section) ----
fig_htn_per1m              <- load_sl("sl_fig_htn_per1m.rds")
fig_tfa_per1m              <- load_sl("sl_fig_tfa_per1m.rds")
fig_combined_per1m         <- load_sl("sl_fig_combined_per1m.rds")
tbl_total_lives_saved      <- load_sl("sl_tbl_total_lives_saved.rds")
tbl_grand_total_lives_saved <- load_sl("sl_tbl_grand_total_lives_saved.rds")
fig_global_cumulative      <- load_sl("sl_fig_global_cumulative.rds")

# ---- New slide artefacts (Bangladesh section) ----
tbl_bangladesh_summary     <- load_sl("sl_tbl_bangladesh_summary.rds")
fig_bangladesh_cumulative  <- load_sl("sl_fig_bangladesh_cumulative.rds")

# ---- PNG paths for pre-rendered figures -------------------------------------
png_htn_program   <- paste0(wd_outp, "assessment_policy_htn_program.png")
png_enrollment    <- paste0(wd_outp, "assessment_enrollment_trend.png")
png_disease_trend <- paste0(wd_outp, "assessment_disease_trend.png")
png_asmr_cause    <- paste0(wd_outp, "assessment_asmr_cause.png")
png_asmr_int      <- paste0(wd_outp, "assessment_asmr_interventions.png")
png_deaths_int    <- paste0(wd_outp, "assessment_fig1_deaths_intervention.png")
png_deaths_cause  <- paste0(wd_outp, "assessment_fig2_deaths_cause.png")
png_annual        <- paste0(wd_outp, "assessment_fig3_annual_averted.png")
png_cum_averted   <- paste0(wd_outp, "assessment_fig4_cum_averted.png")
png_demo_age      <- paste0(wd_outp, "assessment_fig5_demo_age.png")
png_top_countries <- paste0(wd_outp, "assessment_fig5_top_countries.png")
png_map           <- paste0(wd_outp, "assessment_fig6_map.png")

# ---- Display helpers --------------------------------------------------------
fmt_M <- function(x, digits = 1)
  formatC(round(x / 1e6, digits), format = "f", digits = digits, big.mark = ",")

int_order <- c(
  "HTN Control (Program)",
  "TFA Elimination",
  "HTN Control + TFA Elimination"
)

int_cols <- c(
  "HTN Control (Program)"         = "#1f78b4",
  "TFA Elimination"               = "#e31a1c",
  "HTN Control + TFA Elimination" = "#6a3d9a",
  "Counterfactual"                = "#999999"
)

knitr::opts_chunk$set(
  echo      = FALSE,
  message   = FALSE,
  warning   = FALSE,
  fig.align = "center",
  fig.height = 3.6,
  fig.width  = 8.5,
  out.width  = "\\linewidth",
  dev = "cairo_pdf"
)
```

<!-- ======================================================================
     1. TITLE SLIDE (auto-generated from YAML)
     ====================================================================== -->

<!-- ======================================================================
     2. OPENING IMPACT SLIDE
     ====================================================================== -->

## `r if (!is.na(total_combined)) fmt_M(total_combined) else "12"` Million Lives Saved

\justifying

Across `r n_countries` countries, RTSL cardiovascular health programs are projected to save `r if (!is.na(total_combined)) fmt_M(total_combined) else "12"` million lives by 2050.

\medskip

That is approximately **`r if (!is.na(total_combined)) formatC(round(total_combined / (33 * 365)), format = "d", big.mark = ",") else "1,000"` deaths averted every day** for over three decades --- the equivalent of preventing a major disaster, quietly, every single day.

\medskip

\begin{exampleblock}{What This Means}
Sustained investment in hypertension control and trans-fat elimination is delivering one of the largest projected mortality reductions in global cardiovascular health.
\end{exampleblock}

<!-- ======================================================================
     3. WHAT RTSL DOES
     ====================================================================== -->

## What RTSL Does

\justifying

Resolve to Save Lives partners with governments in `r n_countries` countries to deliver two proven, scalable cardiovascular health interventions:

\medskip

- **Hypertension control** --- population-level screening, treatment, and follow-up programs reaching millions of patients since 2017
- **Trans-fat elimination** --- technical assistance to enact and implement best-practice policies that remove industrial trans fats from the food supply
- **Targeting the highest-risk populations** with the most cost-effective interventions available

<!-- ======================================================================
     A5. RTSL HTN PROGRAM ENROLLMENT AND CONTROLLED POPULATION OVER TIME
     ====================================================================== -->

## RTSL Hypertension Program Enrollment and Controlled Population Over Time

```{r fig-enrollment-control-trend, echo=FALSE, out.width="95%", fig.align="center"}
if (file.exists(png_enrollment)) {
  knitr::include_graphics(png_enrollment)
} else {
  plot.new()
  text(
    0.5, 0.5,
    "PLACEHOLDER: Enrollment and controlled population trend\n(knit report.Rmd to generate)",
    cex = 1.2
  )
}
```
\footnotesize RTSL hypertension program enrollment and controlled population over time across all countries.


<!-- ======================================================================
     5. GLOBAL RESULTS
     ====================================================================== -->

# Global Impact

## Cumulative Deaths Averted Over Time --- All Countries

```{r fig-global-cumulative, echo=FALSE, fig.height=3.7}
if (!is.null(fig_global_cumulative)) {
  global_htn <- attr(fig_global_cumulative, "global_cum_htn")
  global_tfa <- attr(fig_global_cumulative, "global_cum_tfa")
  
  global_htn_fmt <- formatC(round(global_htn), format = "d", big.mark = ",")
  global_tfa_fmt <- formatC(round(global_tfa), format = "d", big.mark = ",")
  global_total_fmt <- formatC(round(global_htn + global_tfa), format = "d", big.mark = ",")

  print(fig_global_cumulative)
} else {
  plot.new()
  text(0.5, 0.5, "PLACEHOLDER: Global cumulative chart\n(knit report.Rmd to generate)", cex = 1.2)
  global_htn <- NA_real_
  global_tfa <- NA_real_
  global_htn_fmt <- "N/A"
  global_tfa_fmt <- "N/A"
  global_total_fmt <- "N/A"
}
```


\begin{exampleblock}{By 2050}
\small
`r if (exists("global_htn_fmt")) global_htn_fmt else "N/A"` cumulative deaths averted from HTN control, 
`r if (exists("global_tfa_fmt")) global_tfa_fmt else "N/A"` from TFA elimination, for a combined total of 
`r if (exists("global_htn") & exists("global_tfa")) scales::comma(global_htn + global_tfa) else "N/A"` deaths averted globally.
\end{exampleblock}

## Top Contributing Countries

```{r fig-top-countries, out.width="95%"}
if (file.exists(png_top_countries)) {
  knitr::include_graphics(png_top_countries)
} else {
  plot.new()
  text(0.5, 0.5, "PLACEHOLDER: Top countries\n(knit report.Rmd to generate)", cex = 1.2)
}
```

## Geographic Distribution of Impact

```{r fig-map, out.width="95%"}
if (file.exists(png_map)) {
  knitr::include_graphics(png_map)
} else {
  plot.new()
  text(0.5, 0.5, "PLACEHOLDER: Geographic map\n(knit report.Rmd to generate)", cex = 1.2)
}
```

## Combined Impact: Deaths Averted per 1 Million People

```{r fig-combined-per1m, fig.height=3.6}
if (!is.null(fig_combined_per1m)) {
  print(fig_combined_per1m)
} else {
  plot.new()
  text(0.5, 0.5, "PLACEHOLDER: Combined per-1M results\n(knit report.Rmd to generate)", cex = 1.2)
}
```

\begin{alertblock}{Why Per-Capita Only}
Totals are not shown. Hypertension and TFA interventions have different target populations; per-capita rates enable valid comparison.
\end{alertblock}

<!-- ======================================================================
     4. BANGLADESH CASE EXAMPLE
     ====================================================================== -->

# Impact on the Ground: Bangladesh

## Bangladesh: Cumulative Impact Summary

```{r bangladesh-summary-table, echo=FALSE, results="asis"}
if (!is.null(tbl_bangladesh_summary)) {
  total_bd_val <- attr(tbl_bangladesh_summary, "total_bd_averted")
  total_bd_fmt <- if (!is.null(total_bd_val)) formatC(round(total_bd_val), format = "d", big.mark = ",") else "N/A"

  kable(tbl_bangladesh_summary, format = "latex", booktabs = TRUE,
        caption = "Bangladesh: cumulative program-attributable impact by intervention") |>
    kable_styling(latex_options = c("hold_position"), font_size = 9) |>
    column_spec(3, bold = TRUE)
} else if (!is.null(sl_bangladesh)) {
  kable(sl_bangladesh$summary,
        col.names = c("Scenario", "Deaths Averted (2017--2050)"),
        format = "latex", booktabs = TRUE,
        caption = "Bangladesh: program-attributable deaths averted") |>
    kable_styling(latex_options = c("hold_position"), font_size = 9)
} else {
  cat("\\textit{PLACEHOLDER: Bangladesh summary not yet available. Knit report.Rmd to generate.}")
}
```

\justifying

Bangladesh has been an RTSL hypertension control partner since the program's earliest years. With a large population facing a rising CVD burden, it demonstrates what systematic scale-up can achieve in a low- and middle-income setting.

\footnotesize HTN totals reflect program population; TFA totals reflect national population.

## Bangladesh: Cumulative Deaths Averted Over Time

```{r bangladesh-cumulative-chart, echo=FALSE, fig.height=3.4}
if (!is.null(fig_bangladesh_cumulative)) {
  total_bd_chart <- attr(fig_bangladesh_cumulative, "total_bd")
  total_bd_chart_fmt <- if (!is.null(total_bd_chart)) formatC(round(total_bd_chart), format = "d", big.mark = ",") else "N/A"
  print(fig_bangladesh_cumulative)
} else {
  plot.new()
  text(0.5, 0.5, "PLACEHOLDER: Bangladesh cumulative chart\n(knit report.Rmd to generate)", cex = 1.2)
  total_bd_chart_fmt <- "N/A"
}
```

\begin{exampleblock}{Bangladesh Impact}
Bangladesh: `r if (exists("total_bd_chart_fmt")) total_bd_chart_fmt else "N/A"` total deaths averted from 2017--2050 across both interventions.
\end{exampleblock}

\footnotesize Cumulative deaths averted by intervention type. HTN figures reflect program population; TFA figures reflect national population.

<!-- ======================================================================
     6. HOW WE KNOW THIS
     ====================================================================== -->

# How We Know This

\justifying

- **Grounded in observed data**: Program enrollment, blood pressure control rates, and policy timelines come directly from RTSL program monitoring (2017--2025)
- **Best available epidemiologic evidence**: Baseline disease rates from the Global Burden of Disease 2023; treatment effects from published meta-analyses and clinical trials
- **Country-by-country simulation**: Each of the `r n_countries` countries is modeled individually using its own demographic, epidemiologic, and program data --- projected through 2050
<!-- - **Conservative by design**: The counterfactual assumes zero program effect without RTSL, meaning these projections represent a lower bound on the true impact of the programs -->

<!-- ======================================================================
     7. KEY TAKEAWAYS
     ====================================================================== -->

# Key Takeaways

\justifying

\begin{enumerate}
  \item \textbf{Proven at scale}: RTSL hypertension control programs are the primary driver of projected mortality reductions, operating across `r n_countries` countries and reaching millions of patients
  \item \textbf{Policy impact is real}: Trans-fat elimination policies supported by RTSL are projected to prevent hundreds of thousands of additional deaths from heart disease
  \item \textbf{Scale amplifies returns}: The countries with the largest programs --- India, Bangladesh, China --- contribute the most lives saved, demonstrating the value of continued expansion
\end{enumerate}
<!--  \item \textbf{Conservative estimates}: Because the model assumes zero effect without RTSL, the actual impact is likely even greater than these projections suggest-->

\begin{exampleblock}{Bottom Line}
RTSL programs are projected to avert `r if (!is.na(total_combined)) fmt_M(total_combined) else "X"` million CVD deaths (`r study_years`) across `r n_countries` countries. These are not aspirational targets --- they are grounded in observed program data and available evidence.
\end{exampleblock}

<!-- <!-- ====================================================================== -->
<!--      8. CLOSING DONOR ASK -->
<!--      ====================================================================== --> -->

<!-- # Why Continued Investment Matters -->

<!-- \justifying -->

<!-- \begin{enumerate} -->
<!--   \item \textbf{Programs are working}: Hypertension control rates have risen steadily across RTSL-supported countries since 2017. Sustained funding keeps these gains on track. -->
<!--   \item \textbf{The returns grow over time}: Cardiovascular interventions compound --- every additional year of program operation translates into accelerating lives saved through 2050. -->
<!--   \item \textbf{Disruption costs lives}: If programs stall or lose support, the projected `r if (!is.na(total_combined)) fmt_M(total_combined) else "12"` million deaths averted will not materialize. The counterfactual is not hypothetical --- it is the world without these programs. -->
<!--   \item \textbf{These are conservative projections}: The modeling assumes zero benefit without RTSL. The true value of continued investment is almost certainly higher. -->
<!-- \end{enumerate} -->

<!-- \begin{exampleblock}{The Ask} -->
<!-- Continued and expanded investment in RTSL cardiovascular health programs is among the highest-impact opportunities in global health today. The evidence supports it. The programs are delivering. The lives saved are real. -->
<!-- \end{exampleblock} -->

<!-- ======================================================================
     APPENDIX
     ====================================================================== -->

---

# Appendix

<!-- ======================================================================
     A1. AIMS & RESEARCH QUESTIONS
     ====================================================================== -->

## Central Research Question

\begin{block}{Research Question}
What is the projected health impact --- in terms of \textbf{CVD deaths averted} --- of RTSL cardiovascular health programs operating in `r n_countries` countries from 2017 to 2025, projected to 2050?
\end{block}

\medskip

\begin{enumerate}
  \item How many CVD deaths are attributable to RTSL \textbf{hypertension control programs} (2017--2050)?
  \item What is the additional mortality impact of \textbf{TFA elimination} policies supported by RTSL?
  \item What is the \textbf{combined} program-attributable impact of HTN control and TFA elimination?
  \item How do impacts vary across \textbf{countries}, \textbf{age groups}, and \textbf{causes of death}?
\end{enumerate}

<!-- ======================================================================
     A2. DATA SOURCES & KEY VARIABLES
     ====================================================================== -->

## Data Sources and Key Variables

```{r data-sources-table}
ds <- data.table(
  Source    = c("GBD 2023 (IHME)", "GBD 2019 (IHME)", "Ettehad et al.\\ 2016",
                "UNWPP 2024 (UN)", "HTN Global Tracker", "TFA Policy Registry"),
  Variables = c(
    "Incidence, case fatality, prevalence, background mortality (age/sex/country)",
    "Relative risks per 10 mmHg SBP increase (IHD, stroke, HHD, by age)",
    "Trial effect sizes per BP category; diabetes-stratified",
    "Population projections 2019--2050 (single-year age, sex, country)",
    "Cumulative enrollment, BP controlled, control rates (2017--2025)",
    "Best-practice policy passage/effective dates, RTSL engagement status"
  ),
  Role = c(
    "Baseline transition rates; initial states",
    "BP-bin incidence; normalisation factor $\\alpha$",
    "Treatment effect sizes by BP bin $\\times$ cause",
    "Cohort sizes; rate denominators",
    "Program control rates; enrolled population",
    "TFA policy timing; counterfactual exposure"
  )
)
kable(ds, align = "lll", escape = FALSE, booktabs = TRUE, format = "latex",
      col.names = c("Source", "Variables extracted", "Role in model")) |>
  kable_styling(latex_options = c("striped", "hold_position"),
                font_size = 7, full_width = TRUE) |>
  column_spec(1, bold = TRUE, width = "2.6cm") |>
  column_spec(2, width = "6.5cm") |>
  column_spec(3, width = "3.5cm")
```

<!-- ======================================================================
     A3. COUNTRIES INCLUDED
     ====================================================================== -->

## Countries Included in the Assessment, by Region

```{r tbl-countries-by-region, echo=FALSE, results="asis"}
if (!is.null(tbl_countries_by_region)) {
  kable(tbl_countries_by_region, format = "latex", booktabs = TRUE,
        col.names = c("Region", "Countries"),
        caption = "Countries included in the assessment, by region.") |>
    kable_styling(latex_options = c("striped", "hold_position"),
                  font_size = 8, full_width = TRUE) |>
    column_spec(1, bold = TRUE, width = "3.5cm") |>
    column_spec(2, width = "10cm")
} else {
  cat("\\textit{PLACEHOLDER: Countries by region not yet available. Knit report.Rmd to generate.}")
}
```

- **HTN Module** Excludes Anguilla, British Virgin Islands, St Kitts and Nevis, and Turks and Caicos, as these settings were not included in the 100 Million Lives model.

- **TFA Module** Countries that implemented the ban after the REPLACE initiative.   

<!-- ======================================================================
     A4. OBSERVED AND PROJECTED HTN CONTROL RATES
     ====================================================================== -->

## Observed and Projected Hypertension Control Rates by Country

```{r fig-htn-control-rates, fig.height=4}
if (!is.null(fig_htn_control_rates)) {
  print(fig_htn_control_rates)
} else if (file.exists(png_htn_program)) {
  knitr::include_graphics(png_htn_program)
} else {
  plot.new()
  text(0.5, 0.5, "PLACEHOLDER: HTN control rates\n(knit report.Rmd to generate)", cex = 1.2)
}
```

\note{The "program control rate" is the proportion of enrolled hypertensive patients whose blood pressure is controlled (typically <140/90 mmHg). Observed values (2017--2025) come from the HTN Global Tracker. The projection holds each country's 2024 observed rate constant to 2050 (dashed lines). China, Ecuador, and Bangladesh are highlighted.}

<!-- ======================================================================
     A5. RTSL HTN PROGRAM ENROLLMENT OVER TIME
     ====================================================================== -->

## RTSL Hypertension Program Enrollment Over Time

```{r fig-enrollment-trend, echo=FALSE, out.width="95%", fig.align="center"}
if (file.exists(png_enrollment)) {
  knitr::include_graphics(png_enrollment)
} else {
  plot.new()
  text(0.5, 0.5, "PLACEHOLDER: Enrollment trend\n(knit report.Rmd to generate)", cex = 1.2)
}
```

\footnotesize RTSL hypertension program enrollment over time by country.

<!-- ======================================================================
     A6. ESTIMAND & CAUSAL FRAMEWORK
     ====================================================================== -->

## Estimand and Counterfactual Logic

\begin{block}{Estimand}
The \textbf{program-attributable deaths averted}: the difference in cumulative CVD deaths between the counterfactual scenario (no RTSL program effect) and the intervention scenario (observed program control rates maintained).
\end{block}

\medskip

\justifying

- Each country is simulated **twice** per intervention scenario:
  - **Program run**: observed RTSL control rates applied (2017--2025 data, maintained to 2050)
  - **Counterfactual run**: zero incremental program effect (business-as-usual)
- The **program-attributable impact** = deaths(counterfactual) $-$ deaths(program)

\medskip

\begin{uncertaintyblock}{Counterfactual Assumption}
The counterfactual assumes that without RTSL, HTN control would be 0\% in the objective group and TFA intake  rates would have remained at pre-program levels (zero incremental effect). This is an upper-bound estimate of program impact.
\end{uncertaintyblock}

<!-- ======================================================================
     A7. SCENARIO DEFINITIONS
     ====================================================================== -->

## Scenario Definitions

```{r scenario-table, results='asis'}
dt_scen <- data.table(
  Intervention = c(
    "\\textbf{HTN Control (Program)}",
    "\\textbf{TFA Elimination}",
    "\\textbf{HTN + TFA Combined}"
  ),
  Mechanism = c(
    "Antihypertensive therapy scale-up using observed RTSL program control rates",
    "Complete elimination of industrial TFA in countries with enacted policies",
    "Both interventions applied simultaneously (multiplicative combination)"
  ),
  Key_Param = c(
    "Observed control rates (2017--2025); maintained at 2025 level to 2050",
    "Target: 0\\% energy from TFA; 2-year lag from policy passage to effect",
    "$\\varepsilon_{IR,total} = \\varepsilon_{IR,BP} \\times \\varepsilon_{IR,TFA}$"
  )
)
kable(dt_scen, format = "latex", booktabs = TRUE, escape = FALSE,
      col.names = c("Intervention", "Mechanism", "Key Parameter"),
      caption = NULL) |>
  kable_styling(latex_options = c("hold_position", "scale_down"),
                font_size = 7) |>
  column_spec(1, width = "3.0cm") |>
  column_spec(2, width = "5.5cm") |>
  column_spec(3, width = "4.5cm")
```

<!-- ======================================================================
     A8. MODELING ASSUMPTIONS
     ====================================================================== -->

## Modeling Assumptions

\begin{columns}[T]
\begin{column}{0.47\textwidth}
\textbf{Model structure}
\begin{itemize}\footnotesize
  \item Discrete-time annual step (2017--2050); ages 20--95
  \item Four causes modeled independently: IHD, ischemic stroke, hemorrhagic stroke, HHD
  \item COVID-19 excess mortality applied 2020--2021 only
\end{itemize}

\medskip
\textbf{HTN intervention}
\begin{itemize}\footnotesize
  \item Observed program enrollment \& control rates (2017--2025)
  \item Program population normalised to 1M reference cohort per country-year
  \item Treatment effects from Ettehad meta-analysis, diabetes-weighted
\end{itemize}
\end{column}

\begin{column}{0.50\textwidth}
\textbf{TFA intervention}
\begin{itemize}\footnotesize
  \item Age-specific RRs per 1\% energy from TFA (Wang et al.\ 2016)
  \item Affects IHD case fatality only (no incidence effect)
  \item Policy effective after 2-year implementation lag
\end{itemize}

\medskip
\textbf{Combination rule}
\begin{itemize}\footnotesize
  \item Effect ratios combine multiplicatively across interventions
  \item No assumed interaction between HTN and TFA pathways
\end{itemize}
\end{column}
\end{columns}

<!-- ======================================================================
     A9. KEY LIMITATIONS
     ====================================================================== -->

## Key Limitations

- **Counterfactual assumption**: Business-as-usual assumes zero incremental program effect; actual counterfactual HTN control rates may not be zero
- **Static BP distributions**: No secular trend in population BP distributions; no endogenous updating as program coverage rises
- **Program population normalisation**: Results are scaled to a 1M reference cohort; extrapolation to full national populations requires additional assumptions
- **Control rate maintenance**: 2025 observed rates are held constant to 2050; real-world sustainability is uncertain

<!-- \begin{uncertaintyblock}{Sensitivity Caveats} -->
<!-- The model does not include formal uncertainty quantification (e.g., probabilistic sensitivity analysis). Key parameters --- control rates, effect sizes, TFA exposure levels --- are used at point estimates. -->
<!-- \end{uncertaintyblock} -->

- **TFA data limitations**: Baseline TFA intake is derived from IHME estimates with known measurement challenges
<!-- - **No statin module**: The current assessment does not include a statin therapy intervention (available in the WHO targets model) -->

<!-- ======================================================================
     A10. ADDITIONAL DETAIL TABLES
     ====================================================================== -->

## Countries with TFA Elimination Policies (RTSL-Engaged)

```{r tfa-policy-timeline}
if (!is.null(tbl_tfa_countries)) {
  kable(tbl_tfa_countries, format = "latex", booktabs = TRUE,
        col.names = c("Country", "Year Established", "Year in Effect", "RTSL Engaged")) |>
    kable_styling(latex_options = c("striped", "hold_position"),
                  font_size = 8, full_width = TRUE)
} else {
  cat("\\textit{PLACEHOLDER: TFA policy timeline not yet available. Knit report.Rmd to generate.}")
}
```

\note{This table shows `r if (!is.null(tbl_tfa_countries)) nrow(tbl_tfa_countries) else "N"` countries with enacted TFA elimination policies where RTSL was directly engaged.}

## Hypertension Program Control vs. National Control Rate by Country

```{r tbl-htn-program-vs-country, echo=FALSE, results="asis"}
if (!is.null(tbl_htn_program_vs_country)) {
  underperf <- attr(tbl_htn_program_vs_country, "underperforming_rows")
  k <- kable(tbl_htn_program_vs_country, format = "latex", booktabs = TRUE) |>
    kable_styling(latex_options = c("striped", "hold_position"),
                  font_size = 5, full_width = TRUE)
  if (length(underperf) > 0) {
    k <- k |> row_spec(underperf, background = "#FFDDC1")
  }
  k <- k |> footnote(general = "Highlighted rows: program control rate below national average.")
  k
} else {
  cat("\\textit{PLACEHOLDER: HTN program vs country table not yet available. Knit report.Rmd to generate.}")
}
```

## Total Lives Saved Across All Countries, by Intervention

```{r tbl-grand-total-lives-saved, echo=FALSE, results="asis"}
if (!is.null(tbl_grand_total_lives_saved)) {
  grand_combined_val <- attr(tbl_grand_total_lives_saved, "grand_combined")
  grand_combined_fmt <- if (!is.null(grand_combined_val)) formatC(round(grand_combined_val), format = "d", big.mark = ",") else "N/A"

  kable(tbl_grand_total_lives_saved, format = "latex", booktabs = TRUE) |>
    kable_styling(latex_options = c("hold_position"), font_size = 9) |>
    row_spec(3, bold = TRUE) |>
    footnote(general = "Combined total sums across interventions with different target populations (program pop for HTN; national pop for TFA). Interpret with caution.")
} else {
  cat("\\textit{PLACEHOLDER: Grand total lives saved not yet available. Knit report.Rmd to generate.}")
  grand_combined_fmt <- "N/A"
}
```

\begin{exampleblock}{Grand Total}
Across all RTSL-engaged countries, `r if (exists("grand_combined_fmt")) grand_combined_fmt else "N/A"` total deaths averted from 2017--2050 across both interventions.
\end{exampleblock}

## Total Lives Saved by Country and Intervention

```{r tbl-total-lives-saved, echo=FALSE, results="asis"}
if (!is.null(tbl_total_lives_saved)) {
  kable(tbl_total_lives_saved, format = "latex", booktabs = TRUE) |>
    kable_styling(latex_options = c("striped", "hold_position", "scale_down"),
                  font_size = 7, full_width = TRUE) |>
    footnote(general = "HTN totals reflect program population; TFA totals reflect national population. Combined totals are not shown where populations differ.")
} else {
  cat("\\textit{PLACEHOLDER: Total lives saved table not yet available. Knit report.Rmd to generate.}")
}
```

## Bangladesh: Program vs Overall Population

```{r bangladesh-program-vs-population-short, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
library(knitr)
library(kableExtra)

tab_compare <- data.frame(
  `Age category` = c("≤30", "31–40", "41–50", "51–60", "61–70", "71–80", "81–90", "≥91"),
  `Program %` = c("2.8", "15.4", "28.0", "29.0", "17.9", "5.6", "1.1", "0.2"),
  `Bangladesh %` = c("~26", "~18", "~14", "~11", "~7", "~3", "~1", "~0.2"),
  Interpretation = c("Underrep.", "Slightly underrep.", "Overrep.", "Strongly overrep.",
                     "Overrep.", "Slightly overrep.", "Aligned", "Aligned"),
  check.names = FALSE
)

kbl(tab_compare, booktabs = TRUE, align = "lccc") %>%
  kable_styling(
    latex_options = c("hold_position", "scale_down"),
    font_size = 8
  )

cat("
\\vspace{0.3cm}

{\\footnotesize
\\textbf{Takeaway:} The enrolled population is older than the national population, so the hypertension model should use the program age distribution when estimating impact.
}
")
```

<!-- <!-- ====================================================================== -->
<!--      A11. NEXT STEPS (INTERNAL) -->
<!--      ====================================================================== --> -->

<!-- ## Internal Next Steps -->

<!-- \begin{enumerate} -->
<!--   \item \textbf{Validate country-level inputs} with national program managers --- particularly 2025 control rates and enrollment data -->
<!--   \item \textbf{TFA module} Add counries with RTSL engagement but not with hypertension programs (e.g., Turkey, Ukraine) to the TFA elimination scenario. -->
<!--    \item \textbf{Refine age/sex distributions} Update the model to incorporate actual program-specific age and sex distributions. Current results rely on approximations and are likely \textbf{underestimating impact}, given the observed concentration of enrollment in higher-risk (older) age groups. -->
<!-- \end{enumerate} -->

<!-- <!-- \item \textbf{Integrate statin therapy} for a comprehensive three-intervention assessment --> -->
<!--   <!-- \item \textbf{Add probabilistic sensitivity analysis} to quantify uncertainty around key parameters and provide confidence intervals --> -->

<!-- \medskip -->

<!-- \begin{alertblock}{Decision Needed} -->
<!-- Agree on the reporting unit for external communications: deaths averted per 1M program population vs.\ total national deaths averted. -->
<!-- \end{alertblock} -->

<!-- ======================================================================
     COMMENTED-OUT SLIDES (preserved from original)
     ====================================================================== -->

<!-- ## Deaths Averted by Intervention (All Scenarios) -->

<!-- ```{r fig-deaths-intervention, out.width="95%"} -->
<!-- if (file.exists(png_deaths_int)) { -->
<!--   knitr::include_graphics(png_deaths_int) -->
<!-- } else { -->
<!--   plot.new() -->
<!--   text(0.5, 0.5, "PLACEHOLDER: Deaths averted by intervention\n(knit report.Rmd to generate)", cex = 1.2) -->
<!-- } -->
<!-- ``` -->

<!-- ## Deaths Averted by Cause -->

<!-- ```{r fig-deaths-cause, out.width="95%"} -->
<!-- if (file.exists(png_deaths_cause)) { -->
<!--   knitr::include_graphics(png_deaths_cause) -->
<!-- } else { -->
<!--   plot.new() -->
<!--   text(0.5, 0.5, "PLACEHOLDER: Deaths averted by cause\n(knit report.Rmd to generate)", cex = 1.2) -->
<!-- } -->
<!-- ``` -->

<!-- ## Annual Deaths Averted Over Time -->

<!-- ```{r fig-annual-averted, out.width="95%"} -->
<!-- if (file.exists(png_annual)) { -->
<!--   knitr::include_graphics(png_annual) -->
<!-- } else { -->
<!--   plot.new() -->
<!--   text(0.5, 0.5, "PLACEHOLDER: Annual deaths averted\n(knit report.Rmd to generate)", cex = 1.2) -->
<!-- } -->
<!-- ``` -->

<!-- ## Cumulative Deaths Averted Over Time -->

<!-- ```{r fig-cum-averted, out.width="95%"} -->
<!-- if (file.exists(png_cum_averted)) { -->
<!--   knitr::include_graphics(png_cum_averted) -->
<!-- } else { -->
<!--   plot.new() -->
<!--   text(0.5, 0.5, "PLACEHOLDER: Cumulative deaths averted\n(knit report.Rmd to generate)", cex = 1.2) -->
<!-- } -->
<!-- ``` -->

<!-- ## Deaths Averted by Hypertension Control --- By Country -->

<!-- ```{r fig-htn-per1m, fig.height=3.6} -->
<!-- if (!is.null(fig_htn_per1m)) { -->
<!--   print(fig_htn_per1m) -->
<!-- } else { -->
<!--   plot.new() -->
<!--   text(0.5, 0.5, "PLACEHOLDER: HTN per-1M results\n(knit report.Rmd to generate)", cex = 1.2) -->
<!-- } -->
<!-- ``` -->

<!-- \footnotesize All values are deaths averted per 1 million treated population (2017--2050). -->

<!-- ## Deaths Averted by TFA Elimination --- By Country -->

<!-- ```{r fig-tfa-per1m, fig.height=3.6} -->
<!-- if (!is.null(fig_tfa_per1m)) { -->
<!--   print(fig_tfa_per1m) -->
<!-- } else { -->
<!--   plot.new() -->
<!--   text(0.5, 0.5, "PLACEHOLDER: TFA per-1M results\n(knit report.Rmd to generate)", cex = 1.2) -->
<!-- } -->
<!-- ``` -->

<!-- \footnotesize All values are deaths averted per 1 million population (2017--2050). -->

<!-- ## Model Architecture: Three Health States -->

<!-- ```{r state-diagram, fig.height=2.7} -->
<!-- nodes <- data.table( -->
<!--   x = c(1, 3.2, 5.4), y = c(1, 1, 1), -->
<!--   label = c("Well\n(W)", "Sick\n(K)", "Dead\n(D)"), -->
<!--   fill = c("#d4edda", "#fff3cd", "#f8d7da") -->
<!-- ) -->
<!-- arr_main <- data.table( -->
<!--   x1 = c(1.40, 3.60), y1 = c(1, 1), -->
<!--   x2 = c(2.80, 5.00), y2 = c(1, 1), -->
<!--   lbl = c("IR \u00d7 \u03b5_IR", "CF \u00d7 \u03b5_CF") -->
<!-- ) -->
<!-- arr_bg <- data.table( -->
<!--   x1 = c(1.0, 3.2, 3.0), y1 = c(1.25, 1.25, 1.0), -->
<!--   x2 = c(1.0, 3.2, 5.0), y2 = c(0.62, 0.62, 0.62), -->
<!--   lbl = c("BG.mx", "BG.mx", "BG.mx.all") -->
<!-- ) -->
<!-- ggplot() + -->
<!--   geom_tile(data = nodes, aes(x = x, y = y, fill = fill), -->
<!--             width = 0.75, height = 0.42, color = "grey30", linewidth = 0.8) + -->
<!--   scale_fill_identity() + -->
<!--   geom_text(data = nodes, aes(x = x, y = y, label = label), -->
<!--             size = 4, fontface = "bold") + -->
<!--   geom_segment(data = arr_main, -->
<!--     aes(x = x1, y = y1, xend = x2, yend = y2), -->
<!--     arrow = arrow(length = unit(0.22, "cm"), type = "closed"), -->
<!--     linewidth = 1.2, color = "#2171b5") + -->
<!--   geom_label(data = arr_main, -->
<!--     aes(x = (x1 + x2) / 2, y = y1 + 0.22, label = lbl), -->
<!--     size = 3.2, fill = "white", label.padding = unit(0.12, "cm"), -->
<!--     color = "#2171b5") + -->
<!--   geom_segment(data = arr_bg, -->
<!--     aes(x = x1, y = y1, xend = x2, yend = y2), -->
<!--     arrow = arrow(length = unit(0.16, "cm"), type = "closed"), -->
<!--     linewidth = 0.7, color = "grey55", linetype = "dashed") + -->
<!--   geom_label(data = arr_bg, -->
<!--     aes(x = (x1 + x2) / 2 + 0.2, y = (y1 + y2) / 2, label = lbl), -->
<!--     size = 2.6, fill = "white", color = "grey50", -->
<!--     label.padding = unit(0.08, "cm")) + -->
<!--   scale_x_continuous(limits = c(0.5, 6.0)) + -->
<!--   scale_y_continuous(limits = c(0.28, 1.58)) + -->
<!--   theme_void(base_size = 11) + -->
<!--   labs(title = "Multi-State Transition Model  |  Annual time step  |  Ages 20-95  |  4 CVD causes") -->
<!-- ``` -->

<!-- \small \textbf{Key:} $\varepsilon_{IR}$, $\varepsilon_{CF}$ = multiplicative effect ratios from the intervention module $\cdot$ BG.mx = background mortality $\cdot$ COVID-19 excess mortality applied 2020--2021 only -->

<!-- ## Core Equations -->

<!-- \footnotesize -->

<!-- **State transitions** (age $a \to a+1$, year $t \to t+1$): -->

<!-- $$\text{NewCases}_{a+1,t+1} = W_{a,t} \times IR_{a,t} \times \varepsilon_{IR}$$ -->

<!-- $$K_{a+1,t+1} = K_{a,t}(1 - CF_{a,t}\varepsilon_{CF} - BG.mx - covid.mx) + W_{a,t} \times IR_{a,t}\varepsilon_{IR}$$ -->

<!-- $$D_{a+1,t+1} = K_{a,t} \times CF_{a,t} \times \varepsilon_{CF}$$ -->

<!-- \medskip -->

<!-- \textit{The intervention effect ratios $\varepsilon_{IR}$ and $\varepsilon_{CF}$ are the only quantities that differ between program and counterfactual runs. In the counterfactual, $\varepsilon_{IR} = \varepsilon_{CF} = 1$.} -->

<!-- \medskip -->

<!-- **Combination rule** (multiple interventions): -->

<!-- $$\varepsilon_{IR,\text{total}} = \varepsilon_{IR,\text{BP}} \times \varepsilon_{IR,\text{TFA}}, \quad \varepsilon_{CF,\text{total}} = \varepsilon_{CF,\text{BP}} \times \varepsilon_{CF,\text{TFA}}$$ -->
