---
title: "RTSL Cardiovascular Health Program Impact Assessment"
subtitle: "Projected Health Outcomes 2017--2050"
author: "University of Washington -- Resolve to Save Lives"
institute: "University of Washington"
date: "`r format(Sys.Date(), '%B %Y')`"
output:
  beamer_presentation:
    latex_engine: xelatex
    keep_tex: false
    slide_level: 2
    toc: true
    includes:
      in_header: beamer_preamble.tex
classoption: "aspectratio=169"
bibliography: references.bib
---

<!-- ======================================================================
     SETUP — LOAD ARTEFACTS ONLY (zero recomputation)
     All objects created by report.Rmd and saved to output/ and output/slides/
     ====================================================================== -->

```{r setup, include=FALSE}
library(data.table)
library(ggplot2)
library(knitr)
library(kableExtra)
library(scales)

# ---- Paths ------------------------------------------------------------------
wd       <- normalizePath(file.path(dirname(getwd())), winslash = "/")
# Fallback: hardcode if knitting from docs/
if (!dir.exists(file.path(wd, "output"))) {
 wd <- "C:/Users/wrgar/OneDrive - UW/02Work/ResolveToSaveLives/Assessment/uw-rtsl-assessment/"
}
wd_data  <- paste0(wd, "data/processed/")
wd_outp  <- paste0(wd, "output/")
slides_dir <- paste0(wd_outp, "slides/")

# ---- Load pre-built slide artefacts -----------------------------------------
# These are generated by report.Rmd — knit report.Rmd first if missing.
if (file.exists(paste0(slides_dir, "sl_scalars.rds"))) {
  sl <- readRDS(paste0(slides_dir, "sl_scalars.rds"))
} else {
  sl <- list(n_countries = "N", study_years = "20XX--20XX")
}

n_countries           <- if (!is.null(sl$n_countries))      sl$n_countries      else "N"
study_years           <- if (!is.null(sl$study_years))      sl$study_years      else "20XX--20XX"
total_bp              <- if (!is.null(sl$total_bp))         sl$total_bp         else NA
total_tfa             <- if (!is.null(sl$total_tfa))        sl$total_tfa        else NA
total_combined        <- if (!is.null(sl$total_combined))   sl$total_combined   else NA
lives_per_1M_htn      <- if (!is.null(sl$lives_per_1M_htn))      sl$lives_per_1M_htn      else NA
lives_per_1M_tfa      <- if (!is.null(sl$lives_per_1M_tfa))      sl$lives_per_1M_tfa      else NA
lives_per_1M_combined <- if (!is.null(sl$lives_per_1M_combined)) sl$lives_per_1M_combined else NA

# ---- SL-02 to SL-08: Data tables -------------------------------------------
load_sl <- function(name) {
  f <- paste0(slides_dir, name)
  if (file.exists(f)) readRDS(f) else NULL
}

sl_tab_summary  <- load_sl("sl_tab_summary.rds")
sl_tab_region   <- load_sl("sl_tab_region.rds")
sl_tab_top15    <- load_sl("sl_tab_top15.rds")
sl_dt_cause     <- load_sl("sl_dt_cause.rds")
sl_dt_annual    <- load_sl("sl_dt_annual.rds")
sl_bangladesh   <- load_sl("sl_bangladesh.rds")
sl_tab_scenario <- load_sl("sl_tab_scenario.rds")

# ---- PNG paths for pre-rendered figures -------------------------------------
png_htn_program   <- paste0(wd_outp, "assessment_policy_htn_program.png")
png_enrollment    <- paste0(wd_outp, "assessment_enrollment_trend.png")
png_disease_trend <- paste0(wd_outp, "assessment_disease_trend.png")
png_asmr_cause    <- paste0(wd_outp, "assessment_asmr_cause.png")
png_asmr_int      <- paste0(wd_outp, "assessment_asmr_interventions.png")
png_deaths_int    <- paste0(wd_outp, "assessment_fig1_deaths_intervention.png")
png_deaths_cause  <- paste0(wd_outp, "assessment_fig2_deaths_cause.png")
png_annual        <- paste0(wd_outp, "assessment_fig3_annual_averted.png")
png_cum_averted   <- paste0(wd_outp, "assessment_fig4_cum_averted.png")
png_demo_age      <- paste0(wd_outp, "assessment_fig5_demo_age.png")
png_top_countries <- paste0(wd_outp, "assessment_fig5_top_countries.png")
png_map           <- paste0(wd_outp, "assessment_fig6_map.png")

# ---- Display helpers --------------------------------------------------------
fmt_M <- function(x, digits = 1)
  formatC(round(x / 1e6, digits), format = "f", digits = digits, big.mark = ",")

int_order <- c(
  "HTN Control (Program)",
  "TFA Elimination",
  "HTN Control + TFA Elimination"
)

int_cols <- c(
  "HTN Control (Program)"         = "#1f78b4",
  "TFA Elimination"               = "#e31a1c",
  "HTN Control + TFA Elimination" = "#6a3d9a",
  "Counterfactual"                = "#999999"
)

knitr::opts_chunk$set(
  echo      = FALSE,
  message   = FALSE,
  warning   = FALSE,
  fig.align = "center",
  fig.height = 3.6,
  fig.width  = 8.5,
  out.width  = "\\linewidth",
  dev = "cairo_pdf"
)
```

<!-- ======================================================================
     1. TITLE SLIDE (auto-generated from YAML)
     ====================================================================== -->

<!-- ======================================================================
     2. MOTIVATION & PROBLEM STATEMENT
     ====================================================================== -->

# Motivation & Problem Statement

## The Global CVD Burden

\justifying

- Cardiovascular disease (CVD) is the leading cause of death worldwide, claiming an estimated 17.9 million lives annually
- Over 75\% of CVD deaths occur in low- and middle-income countries
- Hypertension is the single largest modifiable risk factor, affecting over 1.3 billion people globally
- Industrial trans-fatty acids (TFA) in the food supply remain a significant and preventable contributor to ischemic heart disease mortality

\begin{exampleblock}{Key Takeaway}
Scalable, evidence-based interventions --- hypertension control programs and TFA elimination policies --- can avert millions of premature CVD deaths over the next two decades.
\end{exampleblock}

## RTSL Cardiovascular Health Programs

\justifying

- **Resolve to Save Lives (RTSL)** supports governments in `r n_countries` countries to implement proven cardiovascular health interventions
- Two primary program pillars:
  - **Hypertension control**: Population-level screening, treatment, and follow-up programs operating since 2017
  - **Trans-fat elimination**: Technical assistance for best-practice TFA policies (mandatory limits, labeling, reformulation)
- Programs target the populations at highest risk with the most cost-effective interventions

<!-- ======================================================================
     3. AIMS & RESEARCH QUESTIONS
     ====================================================================== -->

# Aims & Research Questions

## Central Research Question

\begin{block}{Research Question}
What is the projected health impact --- in terms of \textbf{CVD deaths averted} --- of RTSL cardiovascular health programs operating in `r n_countries` countries from 2017 to 2024, projected to 2050?
\end{block}

\medskip

\begin{enumerate}
  \item How many CVD deaths are attributable to RTSL \textbf{hypertension control programs} (2017--2050)?
  \item What is the additional mortality impact of \textbf{TFA elimination} policies supported by RTSL?
  \item What is the \textbf{combined} program-attributable impact of HTN control and TFA elimination?
  \item How do impacts vary across \textbf{countries}, \textbf{age groups}, and \textbf{causes of death}?
\end{enumerate}

\begin{alertblock}{Open Question}
Should the primary metric be deaths averted, DALYs, or life-years gained?
\end{alertblock}

<!-- ======================================================================
     4. DATA SOURCES & KEY VARIABLES
     ====================================================================== -->

# Data Sources & Key Variables

## Data Sources and Key Variables

```{r data-sources-table}
ds <- data.table(
  Source    = c("GBD 2023 (IHME)", "GBD 2019 (IHME)", "Ettehad et al.\\ 2016",
                "UNWPP 2024 (UN)", "HTN Global Tracker", "TFA Policy Registry"),
  Variables = c(
    "Incidence, case fatality, prevalence, background mortality (age/sex/country)",
    "Relative risks per 10 mmHg SBP increase (IHD, stroke, HHD, by age)",
    "Trial effect sizes per BP category; diabetes-stratified",
    "Population projections 2019--2050 (single-year age, sex, country)",
    "Cumulative enrollment, BP controlled, control rates (2017--2025)",
    "Best-practice policy passage/effective dates, RTSL engagement status"
  ),
  Role = c(
    "Baseline transition rates; initial states",
    "BP-bin incidence; normalisation factor $\\alpha$",
    "Treatment effect sizes by BP bin $\\times$ cause",
    "Cohort sizes; rate denominators",
    "Program control rates; enrolled population",
    "TFA policy timing; counterfactual exposure"
  )
)
kable(ds, align = "lll", escape = FALSE, booktabs = TRUE, format = "latex",
      col.names = c("Source", "Variables extracted", "Role in model")) |>
  kable_styling(latex_options = c("striped", "hold_position"),
                font_size = 7, full_width = TRUE) |>
  column_spec(1, bold = TRUE, width = "2.6cm") |>
  column_spec(2, width = "6.5cm") |>
  column_spec(3, width = "3.5cm")
```

<!-- ======================================================================
     5. ESTIMAND, CAUSAL FRAMEWORK & COUNTERFACTUAL IMPUTATION
     ====================================================================== -->

# Estimand & Causal Framework

## Estimand and Counterfactual Logic

\begin{block}{Estimand}
The \textbf{program-attributable deaths averted}: the difference in cumulative CVD deaths between the counterfactual scenario (no RTSL program effect) and the intervention scenario (observed program control rates maintained).
\end{block}

\medskip

\justifying

- Each country is simulated **twice** per intervention scenario:
  - **Program run**: observed RTSL control rates applied (2017--2024 data, maintained to 2050)
  - **Counterfactual run**: zero incremental program effect (business-as-usual)
- The **program-attributable impact** = deaths(counterfactual) $-$ deaths(program)

\medskip

\begin{uncertaintyblock}{Counterfactual Assumption}
The counterfactual assumes that without RTSL, HTN control rates would have remained at pre-program levels (zero incremental effect). This is a conservative upper-bound estimate of program impact.
\end{uncertaintyblock}

<!-- ======================================================================
     6. SCENARIOS & ASSUMPTIONS
     ====================================================================== -->

# Scenarios & Assumptions

## Scenario Definitions

```{r scenario-table, results='asis'}
dt_scen <- data.table(
  Intervention = c(
    "\\textbf{HTN Control (Program)}",
    "\\textbf{TFA Elimination}",
    "\\textbf{HTN + TFA Combined}"
  ),
  Mechanism = c(
    "Antihypertensive therapy scale-up using observed RTSL program control rates",
    "Complete elimination of industrial TFA in countries with enacted policies",
    "Both interventions applied simultaneously (multiplicative combination)"
  ),
  Key_Param = c(
    "Observed control rates (2017--2024); maintained at 2024 level to 2050",
    "Target: 0\\% energy from TFA; 2-year lag from policy passage to effect",
    "$\\varepsilon_{IR,total} = \\varepsilon_{IR,BP} \\times \\varepsilon_{IR,TFA}$"
  )
)
kable(dt_scen, format = "latex", booktabs = TRUE, escape = FALSE,
      col.names = c("Intervention", "Mechanism", "Key Parameter"),
      caption = NULL) |>
  kable_styling(latex_options = c("hold_position", "scale_down"),
                font_size = 7) |>
  column_spec(1, width = "3.0cm") |>
  column_spec(2, width = "5.5cm") |>
  column_spec(3, width = "4.5cm")
```

## Modeling Assumptions

\begin{columns}[T]
\begin{column}{0.47\textwidth}
\textbf{Model structure}
\begin{itemize}\footnotesize
  \item Discrete-time annual step (2017--2050); ages 20--95
  \item Four causes modeled independently: IHD, ischemic stroke, hemorrhagic stroke, HHD
  \item COVID-19 excess mortality applied 2020--2021 only
\end{itemize}

\medskip
\textbf{HTN intervention}
\begin{itemize}\footnotesize
  \item Observed program enrollment \& control rates (2017--2024)
  \item Program population normalised to 1M reference cohort per country-year
  \item Treatment effects from Ettehad meta-analysis, diabetes-weighted
\end{itemize}
\end{column}

\begin{column}{0.50\textwidth}
\textbf{TFA intervention}
\begin{itemize}\footnotesize
  \item Age-specific RRs per 1\% energy from TFA (Wang et al.\ 2016)
  \item Affects IHD case fatality only (no incidence effect)
  \item Policy effective after 2-year implementation lag
\end{itemize}

\medskip
\textbf{Combination rule}
\begin{itemize}\footnotesize
  \item Effect ratios combine multiplicatively across interventions
  \item No assumed interaction between HTN and TFA pathways
\end{itemize}
\end{column}
\end{columns}

<!-- ======================================================================
     7. MATHEMATICAL MODEL
     ====================================================================== -->

# Mathematical Model

## Model Architecture: Three Health States

```{r state-diagram, fig.height=2.7}
nodes <- data.table(
  x = c(1, 3.2, 5.4), y = c(1, 1, 1),
  label = c("Well\n(W)", "Sick\n(K)", "Dead\n(D)"),
  fill = c("#d4edda", "#fff3cd", "#f8d7da")
)
arr_main <- data.table(
  x1 = c(1.40, 3.60), y1 = c(1, 1),
  x2 = c(2.80, 5.00), y2 = c(1, 1),
  lbl = c("IR \u00d7 \u03b5_IR", "CF \u00d7 \u03b5_CF")
)
arr_bg <- data.table(
  x1 = c(1.0, 3.2, 3.0), y1 = c(1.25, 1.25, 1.0),
  x2 = c(1.0, 3.2, 5.0), y2 = c(0.62, 0.62, 0.62),
  lbl = c("BG.mx", "BG.mx", "BG.mx.all")
)
ggplot() +
  geom_tile(data = nodes, aes(x = x, y = y, fill = fill),
            width = 0.75, height = 0.42, color = "grey30", linewidth = 0.8) +
  scale_fill_identity() +
  geom_text(data = nodes, aes(x = x, y = y, label = label),
            size = 4, fontface = "bold") +
  geom_segment(data = arr_main,
    aes(x = x1, y = y1, xend = x2, yend = y2),
    arrow = arrow(length = unit(0.22, "cm"), type = "closed"),
    linewidth = 1.2, color = "#2171b5") +
  geom_label(data = arr_main,
    aes(x = (x1 + x2) / 2, y = y1 + 0.22, label = lbl),
    size = 3.2, fill = "white", label.padding = unit(0.12, "cm"),
    color = "#2171b5") +
  geom_segment(data = arr_bg,
    aes(x = x1, y = y1, xend = x2, yend = y2),
    arrow = arrow(length = unit(0.16, "cm"), type = "closed"),
    linewidth = 0.7, color = "grey55", linetype = "dashed") +
  geom_label(data = arr_bg,
    aes(x = (x1 + x2) / 2 + 0.2, y = (y1 + y2) / 2, label = lbl),
    size = 2.6, fill = "white", color = "grey50",
    label.padding = unit(0.08, "cm")) +
  scale_x_continuous(limits = c(0.5, 6.0)) +
  scale_y_continuous(limits = c(0.28, 1.58)) +
  theme_void(base_size = 11) +
  labs(title = "Multi-State Transition Model  |  Annual time step  |  Ages 20-95  |  4 CVD causes")
```

\small \textbf{Key:} $\varepsilon_{IR}$, $\varepsilon_{CF}$ = multiplicative effect ratios from the intervention module $\cdot$ BG.mx = background mortality $\cdot$ COVID-19 excess mortality applied 2020--2021 only

## Core Equations

\footnotesize

**State transitions** (age $a \to a+1$, year $t \to t+1$):

$$\text{NewCases}_{a+1,t+1} = W_{a,t} \times IR_{a,t} \times \varepsilon_{IR}$$

$$K_{a+1,t+1} = K_{a,t}(1 - CF_{a,t}\varepsilon_{CF} - BG.mx - covid.mx) + W_{a,t} \times IR_{a,t}\varepsilon_{IR}$$

$$D_{a+1,t+1} = K_{a,t} \times CF_{a,t} \times \varepsilon_{CF}$$

\medskip

\textit{The intervention effect ratios $\varepsilon_{IR}$ and $\varepsilon_{CF}$ are the only quantities that differ between program and counterfactual runs. In the counterfactual, $\varepsilon_{IR} = \varepsilon_{CF} = 1$.}

\medskip

**Combination rule** (multiple interventions):

$$\varepsilon_{IR,\text{total}} = \varepsilon_{IR,\text{BP}} \times \varepsilon_{IR,\text{TFA}}, \quad \varepsilon_{CF,\text{total}} = \varepsilon_{CF,\text{BP}} \times \varepsilon_{CF,\text{TFA}}$$

<!-- ======================================================================
     8. RESULTS
     ====================================================================== -->

# Results

## Deaths Averted by Intervention

```{r fig-deaths-intervention, out.width="95%"}
if (file.exists(png_deaths_int)) {
  knitr::include_graphics(png_deaths_int)
} else {
  # PLACEHOLDER: Insert cumulative deaths averted bar chart by intervention
  plot.new()
  text(0.5, 0.5, "PLACEHOLDER: Deaths averted by intervention\n(knit report.Rmd to generate)", cex = 1.2)
}
```

## Deaths Averted by Cause

```{r fig-deaths-cause, out.width="95%"}
if (file.exists(png_deaths_cause)) {
  knitr::include_graphics(png_deaths_cause)
} else {
  # PLACEHOLDER: Insert deaths averted broken down by CVD cause
  plot.new()
  text(0.5, 0.5, "PLACEHOLDER: Deaths averted by cause\n(knit report.Rmd to generate)", cex = 1.2)
}
```

## Annual Deaths Averted Over Time

```{r fig-annual-averted, out.width="95%"}
if (file.exists(png_annual)) {
  knitr::include_graphics(png_annual)
} else {
  # PLACEHOLDER: Insert annual deaths averted time series
  plot.new()
  text(0.5, 0.5, "PLACEHOLDER: Annual deaths averted\n(knit report.Rmd to generate)", cex = 1.2)
}
```

## Cumulative Deaths Averted Over Time

```{r fig-cum-averted, out.width="95%"}
if (file.exists(png_cum_averted)) {
  knitr::include_graphics(png_cum_averted)
} else {
  # PLACEHOLDER: Insert cumulative deaths averted time series
  plot.new()
  text(0.5, 0.5, "PLACEHOLDER: Cumulative deaths averted\n(knit report.Rmd to generate)", cex = 1.2)
}
```

## Top Contributing Countries

```{r fig-top-countries, out.width="95%"}
if (file.exists(png_top_countries)) {
  knitr::include_graphics(png_top_countries)
} else {
  # PLACEHOLDER: Insert top countries by deaths averted
  plot.new()
  text(0.5, 0.5, "PLACEHOLDER: Top countries\n(knit report.Rmd to generate)", cex = 1.2)
}
```

## Geographic Distribution of Impact

```{r fig-map, out.width="95%"}
if (file.exists(png_map)) {
  knitr::include_graphics(png_map)
} else {
  # PLACEHOLDER: Insert world map of deaths averted by country
  plot.new()
  text(0.5, 0.5, "PLACEHOLDER: Geographic map\n(knit report.Rmd to generate)", cex = 1.2)
}
```

<!-- ======================================================================
     9. COUNTRY CASE EXAMPLE: BANGLADESH
     ====================================================================== -->

# Country Case Example: Bangladesh

## Bangladesh: Program Context

\justifying

- Bangladesh has been an RTSL hypertension control partner since the early years of the program
- Large population with high CVD burden and rising hypertension prevalence
- Program enrollment and control rates have scaled up progressively since 2017

\begin{exampleblock}{Why Bangladesh?}
Bangladesh demonstrates how a large-population LMIC can achieve measurable CVD mortality reductions through systematic hypertension control at scale.
\end{exampleblock}

## Bangladesh: Key Results

```{r bangladesh-results}
# Pre-computed by report.Rmd -> output/slides/sl_bangladesh.rds
if (!is.null(sl_bangladesh)) {
  # Summary table
  kable(sl_bangladesh$summary,
        col.names = c("Scenario", "Deaths Averted (2017--2050)"),
        format = "latex", booktabs = TRUE,
        caption = "Bangladesh: program-attributable deaths averted") |>
    kable_styling(latex_options = c("hold_position"), font_size = 9)
} else {
  cat("\\textit{PLACEHOLDER: Bangladesh results not yet available.
      Knit report.Rmd to generate sl\\_bangladesh.rds.}")
}
```

## Bangladesh: Deaths Averted by Cause

```{r bangladesh-cause}
if (!is.null(sl_bangladesh) && !is.null(sl_bangladesh$cause)) {
  cause_map_bgd <- c(ihd = "IHD", istroke = "Ischemic stroke",
                      hstroke = "Hemorrhagic stroke", hhd = "Hypertensive HD")
  dt_bgd_cause <- copy(sl_bangladesh$cause)
  dt_bgd_cause[, cause_lbl := cause_map_bgd[cause]]
  dt_bgd_cause[is.na(cause_lbl), cause_lbl := cause]

  kable(dt_bgd_cause[, .(Scenario = scenario, Cause = cause_lbl,
                          `Deaths Averted` = averted)],
        format = "latex", booktabs = TRUE,
        caption = "Bangladesh: deaths averted by scenario and cause") |>
    kable_styling(latex_options = c("hold_position", "scale_down"),
                  font_size = 8)
} else {
  cat("\\textit{PLACEHOLDER: Bangladesh cause-level results not yet available.}")
}
```

<!-- ======================================================================
     10. DISCUSSION & INTERPRETATION
     ====================================================================== -->

# Discussion & Interpretation

## Key Takeaways

\begin{enumerate}
  \item \textbf{Hypertension control programs} are the dominant driver of deaths averted, operating through both reduced disease incidence and improved case fatality
  \item \textbf{TFA elimination} provides additional mortality reductions concentrated in IHD, with effects appearing after a policy implementation lag
  \item \textbf{Combined interventions} yield more than the sum of individual effects due to multiplicative combination of effect ratios
  \item Impact is concentrated in \textbf{older age groups} (60+) and in countries with the \textbf{largest enrolled populations} (India, Bangladesh, China)
  \item Results per 1 million program population provide a \textbf{standardised metric} for cross-country comparison
\end{enumerate}

\begin{exampleblock}{Bottom Line}
RTSL programs are projected to avert `r if (!is.na(total_combined)) fmt_M(total_combined) else "X"` million CVD deaths (`r study_years`) across `r n_countries` countries, with hypertension control as the primary driver and TFA elimination providing complementary gains.
\end{exampleblock}

<!-- ======================================================================
     11. LIMITATIONS
     ====================================================================== -->

# Limitations

## Key Limitations

- **Counterfactual assumption**: Business-as-usual assumes zero incremental program effect; actual counterfactual HTN control rates may not be zero
- **Static BP distributions**: No secular trend in population BP distributions; no endogenous updating as program coverage rises
- **Program population normalisation**: Results are scaled to a 1M reference cohort; extrapolation to full national populations requires additional assumptions
- **Control rate maintenance**: 2024 observed rates are held constant to 2050; real-world sustainability is uncertain

\begin{uncertaintyblock}{Sensitivity Caveats}
The model does not include formal uncertainty quantification (e.g., probabilistic sensitivity analysis). Key parameters --- control rates, effect sizes, TFA exposure levels --- are used at point estimates.
\end{uncertaintyblock}

- **TFA data limitations**: Baseline TFA intake is derived from IHME estimates with known measurement challenges
- **No statin module**: The current assessment does not include a statin therapy intervention (available in the WHO targets model)

<!-- ======================================================================
     12. NEXT STEPS & RECOMMENDATIONS
     ====================================================================== -->

# Next Steps & Recommendations

## Recommended Actions

\begin{enumerate}
  \item \textbf{Validate country-level inputs} with national program managers --- particularly 2024 control rates and enrollment data
  \item \textbf{Add probabilistic sensitivity analysis} to quantify uncertainty around key parameters and provide confidence intervals
  \item \textbf{Extend the TFA module} with updated exposure data and additional country coverage as new policies are enacted
  \item \textbf{Integrate statin therapy} for a comprehensive three-intervention assessment
\end{enumerate}

\medskip

\begin{alertblock}{Decision Needed}
Agree on the reporting unit for external communications: deaths averted per 1M program population vs.\ total national deaths averted.
\end{alertblock}

\begin{exampleblock}{Policy Implication}
These results provide quantitative evidence that RTSL programs are delivering measurable health impact, supporting continued investment and expansion to additional countries.
\end{exampleblock}
