---
title: "RTSL Cardiovascular Health Program Impact Assessment"
subtitle: "Projected Health Outcomes 2017--2050"
author: "University of Washington -- Resolve to Save Lives"
institute: "\\color{red}For validation"
date: "`r format(Sys.Date(), '%B %Y')`"
output:
  beamer_presentation:
    latex_engine: xelatex
    keep_tex: false
    slide_level: 2
    includes:
      in_header: beamer_preamble.tex
classoption: "aspectratio=169"
bibliography: references.bib
---

<!-- ======================================================================
     SETUP — LOAD ARTEFACTS ONLY (zero recomputation)
     All objects created by report.Rmd and saved to output/ and output/slides/
     ====================================================================== -->

```{r setup, include=FALSE}
library(data.table)
library(ggplot2)
library(knitr)
library(kableExtra)
library(scales)

# ---- Paths ------------------------------------------------------------------
#wd       <- normalizePath(file.path(dirname(getwd())), winslash = "/")
# Fallback: hardcode if knitting from docs/
#if (!dir.exists(file.path(wd, "output"))) {
 wd <- "C:/Users/wrgar/OneDrive - UW/02Work/ResolveToSaveLives/Assessment/uw-rtsl-assessment/"
#}
wd_data  <- paste0(wd, "data/processed/")
wd_outp  <- paste0(wd, "output/")
slides_dir <- paste0(wd_outp, "slides/")

# ---- Load pre-built slide artefacts -----------------------------------------
# These are generated by report.Rmd — knit report.Rmd first if missing.
if (file.exists(paste0(slides_dir, "sl_scalars.rds"))) {
  sl <- readRDS(paste0(slides_dir, "sl_scalars.rds"))
} else {
  sl <- list(n_countries = "N", study_years = "20XX--20XX")
}

n_countries           <- if (!is.null(sl$n_countries))      sl$n_countries      else "N"
n_htn_countries       <- if (!is.null(sl$n_htn_countries))  sl$n_htn_countries  else "N"
n_tfa_countries       <- if (!is.null(sl$n_tfa_countries))  sl$n_tfa_countries  else "N"
study_years           <- if (!is.null(sl$study_years))      sl$study_years      else "20XX--20XX"
total_bp              <- if (!is.null(sl$total_bp))         sl$total_bp         else NA
total_tfa             <- if (!is.null(sl$total_tfa))        sl$total_tfa        else NA
total_combined        <- if (!is.null(sl$total_combined))   sl$total_combined   else NA
lives_per_1M_htn      <- if (!is.null(sl$lives_per_1M_htn))      sl$lives_per_1M_htn      else NA
lives_per_1M_tfa      <- if (!is.null(sl$lives_per_1M_tfa))      sl$lives_per_1M_tfa      else NA
lives_per_1M_combined <- if (!is.null(sl$lives_per_1M_combined)) sl$lives_per_1M_combined else NA

# ---- SL-02 to SL-08: Data tables -------------------------------------------
load_sl <- function(name) {
  f <- paste0(slides_dir, name)
  if (file.exists(f)) readRDS(f) else NULL
}

sl_tab_summary  <- load_sl("sl_tab_summary.rds")
sl_tab_region   <- load_sl("sl_tab_region.rds")
sl_tab_top15    <- load_sl("sl_tab_top15.rds")
sl_dt_cause     <- load_sl("sl_dt_cause.rds")
sl_dt_annual    <- load_sl("sl_dt_annual.rds")
sl_bangladesh   <- load_sl("sl_bangladesh.rds")
sl_tab_scenario <- load_sl("sl_tab_scenario.rds")

# ---- New slide artefacts (Data Sources & Key Variables section) ----
fig_htn_control_rates      <- load_sl("sl_fig_htn_control_rates.rds")
tbl_tfa_countries           <- load_sl("sl_tbl_tfa_countries.rds")
tbl_htn_program_vs_country <- load_sl("sl_tbl_htn_program_vs_country.rds")
tbl_countries_by_region    <- load_sl("sl_tbl_countries_by_region.rds")

# ---- New slide artefacts (Results section) ----
fig_htn_per1m              <- load_sl("sl_fig_htn_per1m.rds")
fig_tfa_per1m              <- load_sl("sl_fig_tfa_per1m.rds")
fig_combined_per1m         <- load_sl("sl_fig_combined_per1m.rds")
tbl_total_lives_saved      <- load_sl("sl_tbl_total_lives_saved.rds")
tbl_grand_total_lives_saved <- load_sl("sl_tbl_grand_total_lives_saved.rds")
fig_global_cumulative      <- load_sl("sl_fig_global_cumulative.rds")

# ---- New slide artefacts (Bangladesh section) ----
tbl_bangladesh_summary     <- load_sl("sl_tbl_bangladesh_summary.rds")
fig_bangladesh_cumulative  <- load_sl("sl_fig_bangladesh_cumulative.rds")

# ---- Pareto chart artefacts ----
p_pareto_htn      <- load_sl("sl_pareto_htn.rds")
p_pareto_tfa      <- load_sl("sl_pareto_tfa.rds")
p_pareto_combined <- load_sl("sl_pareto_combined.rds")

# ---- TFA timeline and country map artefacts ----
fig_tfa_timeline  <- load_sl("sl_fig_tfa_timeline.rds")
fig_country_map   <- load_sl("sl_fig_country_map.rds")

# ---- PNG paths for pre-rendered figures -------------------------------------
png_htn_program   <- paste0(wd_outp, "assessment_policy_htn_program.png")
png_enrollment    <- paste0(wd_outp, "assessment_enrollment_trend.png")
png_disease_trend <- paste0(wd_outp, "assessment_disease_trend.png")
png_asmr_cause    <- paste0(wd_outp, "assessment_asmr_cause.png")
png_asmr_int      <- paste0(wd_outp, "assessment_asmr_interventions.png")
png_deaths_int    <- paste0(wd_outp, "assessment_fig1_deaths_intervention.png")
png_deaths_cause  <- paste0(wd_outp, "assessment_fig2_deaths_cause.png")
png_annual        <- paste0(wd_outp, "assessment_fig3_annual_averted.png")
png_cum_averted   <- paste0(wd_outp, "assessment_fig4_cum_averted.png")
png_demo_age      <- paste0(wd_outp, "assessment_fig5_demo_age.png")
png_top_countries <- paste0(wd_outp, "assessment_fig5_top_countries.png")
png_map           <- paste0(wd_outp, "assessment_fig6_map.png")

# ---- Display helpers --------------------------------------------------------
fmt_M <- function(x, digits = 1)
  formatC(round(x / 1e6, digits), format = "f", digits = digits, big.mark = ",")

int_order <- c(
  "HTN Control (Program)",
  "TFA Elimination",
  "HTN Control + TFA Elimination"
)

int_cols <- c(
  "HTN Control (Program)"         = "#1f78b4",
  "TFA Elimination"               = "#e31a1c",
  "HTN Control + TFA Elimination" = "#6a3d9a",
  "Counterfactual"                = "#999999"
)

knitr::opts_chunk$set(
  echo      = FALSE,
  message   = FALSE,
  warning   = FALSE,
  fig.align = "center",
  fig.height = 3.6,
  fig.width  = 8.5,
  out.width  = "\\linewidth",
  dev = "cairo_pdf"
)

# Grand total from table attribute
grand_combined_val <- if (!is.null(tbl_grand_total_lives_saved)) {
  attr(tbl_grand_total_lives_saved, "grand_combined")
} else {
  NA_real_
}

grand_combined_fmt <- if (!is.na(grand_combined_val)) {
  formatC(round(grand_combined_val), format = "d", big.mark = ",")
} else {
  "N/A"
}

grand_combined_m <- if (!is.na(grand_combined_val)) {
  fmt_M(grand_combined_val)
} else {
  "12"
}

grand_daily_fmt <- if (!is.na(grand_combined_val)) {
  formatC(round(grand_combined_val / (33 * 365)), format = "d", big.mark = ",")
} else {
  "1,000"
}

```

<!-- ======================================================================
     1. TITLE SLIDE (auto-generated from YAML)
     ====================================================================== -->

<!-- ======================================================================
     2. OPENING IMPACT SLIDE
     ====================================================================== -->

## `r grand_combined_m` Million Lives Saved

\justifying

Across `r n_countries` countries, RTSL cardiovascular health programs are projected to avert `r grand_combined_m` million deaths from 2017--2050 across hypertension control and trans-fat elimination interventions.


\medskip

That is approximately **`r grand_daily_fmt` deaths averted every day** over more than three decades.

\medskip

\begin{exampleblock}{What This Means}
Sustained investment in hypertension control and trans-fat elimination could deliver one of the largest projected reductions in cardiovascular mortality across RTSL-engaged countries.
\end{exampleblock}

<!-- ======================================================================
     3. WHAT RTSL DOES
     ====================================================================== -->

## What RTSL Does

\justifying

RTSL collaborates with governments and other partners in more than `r n_countries` countries to deliver proven, scalable CVH interventions:

\medskip

- **Hypertension control** --- blood pressure screening, treatment, and follow-up programs reaching millions of patients since 2017
- **Trans-fat elimination** --- technical assistance and advocacy to enact and implement best-practice policies that remove industrial trans fats from the food supply
- **Targeting the highest-risk populations** with the most cost-effective interventions available

## The Global CVD Burden

\justifying

- Cardiovascular disease (CVD) is the leading cause of death worldwide, claiming an estimated 17.9 million lives annually
- Over 75\% of CVD deaths occur in low- and middle-income countries
- Hypertension is the single largest modifiable risk factor, affecting over 1.3 billion people globally
- Industrial trans-fatty acids (TFA) in the food supply remain a significant and preventable contributor to ischemic heart disease mortality

\begin{exampleblock}{Key Takeaway}
Scalable, evidence-based interventions --- hypertension control programs and TFA elimination policies --- can avert millions of premature CVD deaths over the next two decades.
\end{exampleblock}

<!-- ======================================================================
     A5. RTSL HTN PROGRAM ENROLLMENT AND CONTROLLED POPULATION OVER TIME
     ====================================================================== -->

## RTSL Hypertension Program Enrollment and Controlled Population Over Time

```{r fig-enrollment-control-trend, echo=FALSE, out.width="95%", fig.align="center"}
if (file.exists(png_enrollment)) {
  knitr::include_graphics(png_enrollment)
} else {
  plot.new()
  text(
    0.5, 0.5,
    "PLACEHOLDER: Enrollment and controlled population trend\n(knit report.Rmd to generate)",
    cex = 1.2
  )
}
```
\footnotesize Notes: RTSL hypertension program enrollment and control trends across countries. Data from HTN Global Tracker (2017–2025); gaps filled via interpolation/carry-forward. “Program controlled” = enrolled patients with BP <140/90 mmHg. Post-2025, both enrollment and control rates are held constant at 2025 levels for projections. 

<!-- ======================================================================
     TFA TIMELINE AND COUNTRY MAP SLIDES
     ====================================================================== -->

## Trans Fat Elimination: Global Policy Progress

```{r fig-tfa-timeline, echo=FALSE, fig.width=12, fig.height=6, out.width="\\linewidth"}
if (!is.null(fig_tfa_timeline)) {
  print(fig_tfa_timeline)
} else {
  plot.new()
  text(0.5, 0.5, "PLACEHOLDER: TFA timeline chart\n(knit report.Rmd to generate)", cex = 1.2)
}
```

## Global Coverage: HTN and TFA Interventions

```{r fig-country-map, echo=FALSE, fig.width=12, fig.height=6, out.width="\\linewidth"}
if (!is.null(fig_country_map)) {
  print(fig_country_map)
} else {
  plot.new()
  text(0.5, 0.5, "PLACEHOLDER: Country coverage map\n(knit report.Rmd to generate)", cex = 1.2)
}
```

<!-- ======================================================================
     5. GLOBAL RESULTS
     ====================================================================== -->

# Global Impact

## Total Lives Saved Across All Countries, by Intervention

```{r tbl-grand-total-lives-saved, echo=FALSE, results="asis"}
if (!is.null(tbl_grand_total_lives_saved)) {
  grand_combined_val <- attr(tbl_grand_total_lives_saved, "grand_combined")
  grand_combined_fmt <- if (!is.null(grand_combined_val)) formatC(round(grand_combined_val), format = "d", big.mark = ",") else "N/A"

  kable(tbl_grand_total_lives_saved, format = "latex", booktabs = TRUE,
        caption = "All countries: cumulative program-attributable impact by intervention") |>
    kable_styling(latex_options = c("hold_position"), font_size = 9) |>
    column_spec(2, bold = TRUE)
} else {
  cat("\\textit{PLACEHOLDER: Grand total lives saved not yet available. Knit report.Rmd to generate.}")
  grand_combined_fmt <- "N/A"
}
```

\justifying

Across `r n_countries` countries, RTSL cardiovascular health programs are projected to avert `r grand_combined_m` million deaths from 2017--2050. Hypertension control programs operate in `r n_htn_countries` countries, while trans-fat elimination policies cover `r n_tfa_countries` countries --- together representing the full scope of RTSL's global cardiovascular health portfolio.

\footnotesize HTN totals reflect program population; TFA totals reflect national population.


<!-- ======================================================================
     PARETO CHARTS: TOP 20 COUNTRIES
     ====================================================================== -->

## Hypertension Control: Top 20 Countries by Deaths Averted

```{r pareto-htn-slide, echo=FALSE, fig.width=12, fig.height=6, out.width="\\linewidth"}
if (!is.null(p_pareto_htn)) {
  print(p_pareto_htn)
} else {
  plot.new()
  text(0.5, 0.5, "PLACEHOLDER: HTN Pareto chart\n(knit report.Rmd to generate)", cex = 1.2)
}
```

## Trans Fat Elimination: Top 20 Countries by Deaths Averted

```{r pareto-tfa-slide, echo=FALSE, fig.width=12, fig.height=6, out.width="\\linewidth"}
if (!is.null(p_pareto_tfa)) {
  print(p_pareto_tfa)
} else {
  plot.new()
  text(0.5, 0.5, "PLACEHOLDER: TFA Pareto chart\n(knit report.Rmd to generate)", cex = 1.2)
}
```

## Combined Interventions: Top 20 Countries by Deaths Averted

```{r pareto-combined-slide, echo=FALSE, fig.width=12, fig.height=6, out.width="\\linewidth"}
if (!is.null(p_pareto_combined)) {
  print(p_pareto_combined)
} else {
  plot.new()
  text(0.5, 0.5, "PLACEHOLDER: Combined Pareto chart\n(knit report.Rmd to generate)", cex = 1.2)
}
```


<!-- ======================================================================
     6. HOW WE KNOW THIS
     ====================================================================== -->

# How We Know This

\justifying

- **Grounded in observed data**: Program enrollment, blood pressure control rates, and policy timelines come directly from RTSL program monitoring (2017--2025)
- **Best available epidemiologic evidence**: Baseline disease rates from the Global Burden of Disease 2023; treatment effects from published meta-analyses and clinical trials
- **Country-by-country simulation**: Each of the `r n_countries` countries is modeled individually using its own demographic, epidemiologic, and program data --- projected through 2050
<!-- - **Conservative by design**: The counterfactual assumes zero program effect without RTSL, meaning these projections represent a lower bound on the true impact of the programs -->

<!-- ======================================================================
     7. KEY TAKEAWAYS
     ====================================================================== -->

# Key Takeaways

\justifying

\begin{enumerate}
  \item \textbf{Trans-fat elimination delivers the largest population impact}: TFA policies avert the greatest number of deaths due to their population-wide reach, affecting cardiovascular risk across entire countries.

  \item \textbf{Hypertension control delivers scalable program impact}: RTSL-supported hypertension control programs generate substantial mortality reductions among enrolled patients, with total impact increasing as programs expand coverage.

  \item \textbf{Scale and coverage shape total impact}: Countries with large populations and broad implementation --- including India, Bangladesh, and Pakistan --- account for the greatest absolute reductions in mortality.
\end{enumerate}

\begin{exampleblock}{Bottom Line}
Across `r n_countries` countries, RTSL-supported hypertension control programs and trans-fat elimination policies are projected to avert `r grand_combined_m` million deaths from 2017--2050, reflecting the combined impact of population-wide and program-based interventions.
\end{exampleblock}


<!-- ======================================================================
     APPENDIX
     ====================================================================== -->

---

# Appendix

<!-- ======================================================================
     A1. AIMS & RESEARCH QUESTIONS
     ====================================================================== -->

## Central Research Question

\begin{block}{Research Question}
What is the projected health impact --- in terms of \textbf{CVD deaths averted} --- of RTSL cardiovascular health programs operating in `r n_countries` countries from 2017 to 2025, projected to 2050?
\end{block}

\medskip

\begin{enumerate}
  \item How many CVD deaths are attributable to RTSL \textbf{hypertension control programs} (2017--2050)?
  \item What is the additional mortality impact of \textbf{TFA elimination} policies supported by RTSL?
  \item What is the \textbf{combined} program-attributable impact of HTN control and TFA elimination?
  \item How do impacts vary across \textbf{countries}, \textbf{age groups}, and \textbf{causes of death}?
\end{enumerate}

<!-- ======================================================================
     A2. DATA SOURCES & KEY VARIABLES
     ====================================================================== -->

## Data Sources and Key Variables

```{r data-sources-table}
ds <- data.table(
  Source    = c("GBD 2023 (IHME)", "GBD 2019 (IHME)", "Ettehad et al.\\ 2016",
                "UNWPP 2024 (UN)", "HTN Global Tracker", "TFA Policy Registry"),
  Variables = c(
    "Incidence, case fatality, prevalence, background mortality (age/sex/country)",
    "Relative risks per 10 mmHg SBP increase (IHD, stroke, HHD, by age)",
    "Trial effect sizes per BP category; diabetes-stratified",
    "Population projections 2019--2050 (single-year age, sex, country)",
    "Cumulative enrollment, BP controlled, control rates (2017--2025)",
    "Best-practice policy passage/effective dates, RTSL engagement status"
  ),
  Role = c(
    "Baseline transition rates; initial states",
    "BP-bin incidence; normalisation factor $\\alpha$",
    "Treatment effect sizes by BP bin $\\times$ cause",
    "Cohort sizes; rate denominators",
    "Program control rates; enrolled population",
    "TFA policy timing; counterfactual exposure"
  )
)
kable(ds, align = "lll", escape = FALSE, booktabs = TRUE, format = "latex",
      col.names = c("Source", "Variables extracted", "Role in model")) |>
  kable_styling(latex_options = c("striped", "hold_position"),
                font_size = 7, full_width = TRUE) |>
  column_spec(1, bold = TRUE, width = "2.6cm") |>
  column_spec(2, width = "6.5cm") |>
  column_spec(3, width = "3.5cm")
```

<!-- ======================================================================
     A3. COUNTRIES INCLUDED
     ====================================================================== -->

## Countries Included in the Assessment, by Region

```{r tbl-countries-by-region, echo=FALSE, results="asis"}
if (!is.null(tbl_countries_by_region)) {
  kable(tbl_countries_by_region, format = "latex", booktabs = TRUE,
        col.names = c("Region", "Countries"),
        caption = "Countries included in the assessment, by region.") |>
    kable_styling(latex_options = c("striped", "hold_position"),
                  font_size = 8, full_width = TRUE) |>
    column_spec(1, bold = TRUE, width = "3.5cm") |>
    column_spec(2, width = "10cm")
} else {
  cat("\\textit{PLACEHOLDER: Countries by region not yet available. Knit report.Rmd to generate.}")
}
```

- **HTN Module** Excludes Anguilla, British Virgin Islands, St Kitts and Nevis, and Turks and Caicos, as these settings were not included in the 100 Million Lives model.

- **TFA Module** Countries that implemented the ban after the REPLACE initiative.   

<!-- ======================================================================
     A4. OBSERVED AND PROJECTED HTN CONTROL RATES
     ====================================================================== -->

## Observed and Projected Hypertension Control Rates by Country

```{r fig-htn-control-rates, fig.height=4}
if (!is.null(fig_htn_control_rates)) {
  print(fig_htn_control_rates)
} else if (file.exists(png_htn_program)) {
  knitr::include_graphics(png_htn_program)
} else {
  plot.new()
  text(0.5, 0.5, "PLACEHOLDER: HTN control rates\n(knit report.Rmd to generate)", cex = 1.2)
}
```

\note{The "program control rate" is the proportion of enrolled hypertensive patients whose blood pressure is controlled (typically <140/90 mmHg). Observed values (2017--2025) come from the HTN Global Tracker. The projection holds each country's 2024 observed rate constant to 2050 (dashed lines). China, Ecuador, and Bangladesh are highlighted.}

<!-- ======================================================================
     A5. RTSL HTN PROGRAM ENROLLMENT OVER TIME
     ====================================================================== -->

<!-- ======================================================================
     A6. ESTIMAND & CAUSAL FRAMEWORK
     ====================================================================== -->

## Estimand and Counterfactual Logic

\begin{block}{Estimand}
The \textbf{program-attributable deaths averted}: the difference in cumulative CVD deaths between the counterfactual scenario (no RTSL program effect) and the intervention scenario (observed program control rates maintained).
\end{block}

\medskip

\justifying

- Each country is simulated **twice** per intervention scenario:
  - **Program run**: observed RTSL control rates applied (2017--2025 data, maintained to 2050)
  - **Counterfactual run**: zero incremental program effect (business-as-usual)
- The **program-attributable impact** = deaths(counterfactual) $-$ deaths(program)

\medskip

\begin{uncertaintyblock}{Counterfactual Assumption}
The counterfactual assumes that without RTSL, HTN control would be 0\% in the objective group and TFA intake  rates would have remained at pre-program levels (zero incremental effect). This is an upper-bound estimate of program impact.
\end{uncertaintyblock}

<!-- ======================================================================
     A7. SCENARIO DEFINITIONS
     ====================================================================== -->

## Scenario Definitions

```{r scenario-table, results='asis'}
dt_scen <- data.table(
  Intervention = c(
    "\\textbf{HTN Control (Program)}",
    "\\textbf{TFA Elimination}",
    "\\textbf{HTN + TFA Combined}"
  ),
  Mechanism = c(
    "Antihypertensive therapy scale-up using observed RTSL program control rates",
    "Complete elimination of industrial TFA in countries with enacted policies",
    "Both interventions applied simultaneously (multiplicative combination)"
  ),
  Key_Param = c(
    "Observed control rates (2017--2025); maintained at 2025 level to 2050",
    "Target: 0\\% energy from TFA; 2-year lag from policy passage to effect",
    "$\\varepsilon_{IR,total} = \\varepsilon_{IR,BP} \\times \\varepsilon_{IR,TFA}$"
  )
)
kable(dt_scen, format = "latex", booktabs = TRUE, escape = FALSE,
      col.names = c("Intervention", "Mechanism", "Key Parameter"),
      caption = NULL) |>
  kable_styling(latex_options = c("hold_position", "scale_down"),
                font_size = 7) |>
  column_spec(1, width = "3.0cm") |>
  column_spec(2, width = "5.5cm") |>
  column_spec(3, width = "4.5cm")
```

<!-- ======================================================================
     A8. MODELING ASSUMPTIONS
     ====================================================================== -->

## Modeling Assumptions

\begin{columns}[T]
\begin{column}{0.47\textwidth}
\textbf{Model structure}
\begin{itemize}\footnotesize
  \item Discrete-time annual step (2017--2050); ages 20--95
  \item Four causes modeled independently: IHD, ischemic stroke, hemorrhagic stroke, HHD
  \item COVID-19 excess mortality applied 2020--2021 only
\end{itemize}

\medskip
\textbf{HTN intervention}
\begin{itemize}\footnotesize
  \item Observed program enrollment \& control rates (2017--2025)
  \item Program population normalised to 1M reference cohort per country-year
  \item Treatment effects from Ettehad meta-analysis, diabetes-weighted
\end{itemize}
\end{column}

\begin{column}{0.50\textwidth}
\textbf{TFA intervention}
\begin{itemize}\footnotesize
  \item Age-specific RRs per 1\% energy from TFA (Wang et al.\ 2016)
  \item Affects IHD case fatality only (no incidence effect)
  \item Policy effective after 2-year implementation lag
\end{itemize}

\medskip
\textbf{Combination rule}
\begin{itemize}\footnotesize
  \item Effect ratios combine multiplicatively across interventions
  \item No assumed interaction between HTN and TFA pathways
\end{itemize}
\end{column}
\end{columns}

<!-- ======================================================================
     A9. KEY LIMITATIONS
     ====================================================================== -->

## Key Limitations

- **Counterfactual assumption**: Business-as-usual assumes zero incremental program effect; actual counterfactual HTN control rates may not be zero
- **Static BP distributions**: No secular trend in population BP distributions; no endogenous updating as program coverage rises
- **Program population normalisation**: Results are scaled to a 1M reference cohort; extrapolation to full national populations requires additional assumptions
- **Control rate maintenance**: 2025 observed rates are held constant to 2050; real-world sustainability is uncertain

- **TFA data limitations**: Baseline TFA intake is derived from IHME estimates with known measurement challenges

