---
title: "RTSL Cardiovascular Health Program Impact"
subtitle: "Projected health outcomes across RTSL-engaged countries, 2017--2050"
author: "University of Washington -- Resolve to Save Lives"
date: "`r format(Sys.Date(), '%B %Y')`"
output:
  beamer_presentation:
    latex_engine: xelatex
    keep_tex: true
    slide_level: 2
    includes:
      in_header: beamer_preamble.tex
classoption: "aspectratio=169"
bibliography: references.bib
---

<!-- ======================================================================
     SETUP — LOAD ARTEFACTS ONLY (no computation; all objects come from
     report.Rmd). If an artefact is missing, knit report.Rmd first.
     ====================================================================== -->

```{r setup, include=FALSE}
library(data.table)
library(ggplot2)
library(knitr)
library(kableExtra)
library(scales)

# ---- Paths ------------------------------------------------------------------
wd <- "C:/Users/wrgar/OneDrive - UW/02Work/ResolveToSaveLives/Assessment/uw-rtsl-assessment/"
wd_data    <- paste0(wd, "data/processed/")
wd_outp    <- paste0(wd, "output/")
slides_dir <- paste0(wd_outp, "slides/")

load_sl <- function(name) {
  f <- paste0(slides_dir, name)
  if (file.exists(f)) readRDS(f) else NULL
}

# ---- Combined scalars (from report.Rmd pd-combined-scalars chunk) -----------
sl <- load_sl("sl_pd_scalars.rds")
if (is.null(sl)) {
  sl <- list(
    n_countries = "N", n_htn_countries = "N", n_tfa_countries = "N",
    study_years = "20XX-20XX", grand_combined_fmt = "N/A",
    grand_combined_m = "12", daily_averted_fmt = "1,000",
    n_years = 34,
    cvd_deaths_global = "19.2 million",
    cvd_share_noncomm = "approximately one-third",
    htn_adults_global = "1.3 billion",
    htn_uncontrolled_pct = "over 75\\%",
    htn_undiagnosed_share = "nearly half",
    htn_deaths_annual = "10.8 million",
    tfa_deaths_annual = "500,000",
    tfa_protected_pop = "3.4 billion",
    tfa_unprotected_pop = "roughly 5 billion",
    tfa_countries_bpp = "53",
    tfa_countries_none = "over 130",
    rtsl_htn_countries = "30+",
    rtsl_htn_enrolled = "30 million",
    lmic_cvd_share = "over 80\\%",
    premature_cvd_lmic = "over three-quarters"
  )
}

# Shortcut scalars
n_countries        <- sl$n_countries
n_htn_countries    <- sl$n_htn_countries
n_tfa_countries    <- sl$n_tfa_countries
study_years        <- sl$study_years
grand_combined_fmt <- sl$grand_combined_fmt
grand_combined_m   <- sl$grand_combined_m
daily_averted_fmt  <- sl$daily_averted_fmt

# ---- Combined artefacts for results section ---------------------------------
fig_pd_global_cumulative <- load_sl("sl_pd_fig_global_cumulative.rds")
p_pd_pareto_combined     <- load_sl("sl_pd_pareto_combined.rds")
sl_pd_tab_region         <- load_sl("sl_pd_tab_region.rds")
sl_pd_tab_top_countries  <- load_sl("sl_pd_tab_top_countries.rds")

# ---- Pre-rendered (kept) artefacts ------------------------------------------
fig_tfa_timeline  <- load_sl("sl_fig_tfa_timeline.rds")
fig_country_map   <- load_sl("sl_fig_country_map.rds")

png_enrollment    <- paste0(wd_outp, "assessment_enrollment_trend.png")

knitr::opts_chunk$set(
  echo       = FALSE,
  message    = FALSE,
  warning    = FALSE,
  fig.align  = "center",
  fig.height = 3.3,
  fig.width  = 8.3,
  out.width  = "0.92\\linewidth",
  dev        = "cairo_pdf"
)
```

<!-- ======================================================================
     1. TITLE SLIDE (auto-generated from YAML)
     ====================================================================== -->

<!-- ======================================================================
     2. SETTING THE STAGE — 3-column framing slide
     ====================================================================== -->

## Setting the Stage

\begin{columns}[T]
\begin{column}{0.32\textwidth}
\textbf{\color{uwpurple}What is the problem?}
\vspace{0.2em}

\footnotesize
\begin{itemize}
  \item \textbf{Hypertension} and \textbf{industrially-produced trans fat} are two of the most preventable drivers of cardiovascular disease.
  \item Together, they account for millions of premature deaths every year.
\end{itemize}
\scriptsize \textit{Sources: GBD 2023 (IHME); WHO Global Report on Hypertension.}
\end{column}

\begin{column}{0.32\textwidth}
\textbf{\color{uwpurple}Why does it matter?}
\vspace{0.2em}

\footnotesize
\begin{itemize}
  \item CVD causes \textbf{`r sl$cvd_deaths_global`} deaths each year globally --- `r sl$cvd_share_noncomm` of all deaths.
  \item High blood pressure alone drives \textbf{`r sl$htn_deaths_annual`} deaths annually.
  \item Trans fat adds another \textbf{`r sl$tfa_deaths_annual`} IHD deaths.
\end{itemize}
\scriptsize \textit{Sources: Lancet 2025; WHO; GBD 2023.}
\end{column}

\begin{column}{0.32\textwidth}
\textbf{\color{uwpurple}Who is affected?}
\vspace{0.2em}

\footnotesize
\begin{itemize}
  \item \textbf{over 80\%} of CVD deaths occur in low- and middle-income countries.
  \item \textbf{`r sl$premature_cvd_lmic`} of \textit{premature} CVD deaths are in LMICs.
  \item Populations without access to treatment or policy protection bear the greatest burden.
\end{itemize}
\scriptsize \textit{Sources: Lancet 2025; WHO.}
\end{column}
\end{columns}


<!-- ======================================================================
     3. THE CHALLENGE IN NUMBERS
     ====================================================================== -->

## The Challenge in Numbers {.fragile}

\begin{columns}[T]
\begin{column}{0.48\textwidth}
\textbf{\color{uwpurple}Hypertension}
\vspace{0.4em}

{\huge \textbf{`r sl$htn_adults_global`}} \newline
\small adults live with hypertension worldwide

\vspace{0.5em}
{\huge \textbf{over 75\%}} \newline
\small of them are not controlled

\vspace{0.5em}
{\huge \textbf{`r sl$htn_undiagnosed_share`}} \newline
\small are undiagnosed

\vspace{0.8em}
\scriptsize \textit{WHO Global Report on Hypertension (2023/2025 update); resolvetosavelives.org/blood-pressure-control.}
\end{column}

\begin{column}{0.48\textwidth}
\textbf{\color{uwpurple}Trans fat}
\vspace{0.4em}

{\huge \textbf{67}} \newline
\small countries with WHO best-practice TFA policies

\vspace{0.5em}
{\huge \textbf{`r sl$tfa_unprotected_pop`}} \newline
\small people still unprotected

\vspace{0.5em}
{\huge \textbf{`r sl$tfa_deaths_annual`}} \newline
\small IHD deaths every year from industrial trans fat

\vspace{0.8em}
\scriptsize \textit{WHO REPLACE; resolvetosavelives.org/healthier-food.}
\end{column}
\end{columns}

<!-- ======================================================================
     4. OUR APPROACH — 3-column
     ====================================================================== -->
## What RTSL Does

\justifying

RTSL collaborates with governments and other partners in more than `r n_countries` countries to deliver proven, scalable CVH interventions:

\medskip

- **Hypertension control** --- blood pressure screening, treatment, and follow-up programs reaching millions of patients since 2017
- **Trans-fat elimination** --- technical assistance and advocacy to enact and implement best-practice policies that remove industrial trans fats from the food supply
- **Targeting the highest-risk populations** with the most cost-effective interventions available

## Our Approach {.fragile}

\begin{columns}[T]
\begin{column}{0.32\textwidth}
\textbf{\color{uwpurple}What we did}
\vspace{0.2em}

\footnotesize
\begin{itemize}
  \item Built a country-level, multi-state CVD simulation across `r n_countries` countries.
  \item Projected deaths averted from 2017 to 2050 under observed RTSL program and policy trajectories.
  \item Combined HTN control and TFA elimination into one integrated assessment.
\end{itemize}
\end{column}

\begin{column}{0.32\textwidth}
\textbf{\color{uwpurple}Who we worked with}
\vspace{0.2em}

\footnotesize
\begin{itemize}
  \item \textbf{Resolve to Save Lives} --- program, policy, and country teams.
  \item \textbf{Country governments and ministries of health} providing program data.
  \item \textbf{WHO} (REPLACE, HEARTS) policy frameworks and data.
  \item \textbf{IHME / GBD} for baseline epidemiology.
\end{itemize}
\end{column}

\begin{column}{0.32\textwidth}
\textbf{\color{uwpurple}How it was implemented}
\vspace{0.2em}

\footnotesize
\begin{itemize}
  \item Country-by-country modeling using RTSL monitoring data (2017--2025) and UNWPP 2024 projections.
  \item Scenario analysis: program vs. counterfactual, for each intervention.
  \item Results reviewed with RTSL and country partners.
\end{itemize}
\end{column}
\end{columns}

<!-- ======================================================================
     5. HEADLINE IMPACT SLIDE
     ====================================================================== -->

## Headline Impact: `r grand_combined_m` Million Lives Saved {.fragile}

\justifying

\medskip

Across \textbf{`r n_countries` countries}, RTSL cardiovascular health programs are projected to avert \textbf{`r grand_combined_m` deaths} from `r study_years`, combining hypertension control and trans-fat elimination.

\medskip

That is approximately \textbf{`r daily_averted_fmt` deaths averted every day} over more than three decades.

\medskip

\begin{exampleblock}{What This Means}
Sustained investment in hypertension control and trans-fat elimination is projected to deliver one of the largest preventable reductions in cardiovascular mortality across RTSL-engaged countries.
\end{exampleblock}

<!-- <!-- ====================================================================== -->
<!--      6. GLOBAL CUMULATIVE IMPACT CHART (combined only) -->
<!--      ====================================================================== -->

<!-- ## Cumulative Lives Saved Over Time -->

<!-- ```{r fig-global-cumulative-combined, fig.width=11, fig.height=5, out.width="\\linewidth"} -->
<!-- if (!is.null(fig_pd_global_cumulative)) { -->
<!--   print(fig_pd_global_cumulative) -->
<!-- } else { -->
<!--   plot.new() -->
<!--   text(0.5, 0.5, -->
<!--        "PLACEHOLDER: cumulative deaths averted (knit report.Rmd)", -->
<!--        cex = 1.2) -->
<!-- } -->
<!-- ``` -->

<!-- \footnotesize Combined hypertension control and trans-fat elimination, all RTSL-engaged countries, 2017--2050. -->

<!-- <!-- ====================================================================== -->
<!--      7. REGIONAL SUMMARY (combined) -->
<!--      ====================================================================== -->

<!-- ## Impact by Region -->

<!-- ```{r tbl-region-combined, results="asis"} -->
<!-- if (!is.null(sl_pd_tab_region)) { -->
<!--   kable(sl_pd_tab_region, format = "latex", booktabs = TRUE, -->
<!--         align = "lcr", -->
<!--         caption = "Combined deaths averted by region (HTN + TFA, 2017-2050).") |> -->
<!--     kable_styling(latex_options = c("hold_position"), -->
<!--                   font_size = 10, full_width = FALSE) |> -->
<!--     column_spec(1, bold = TRUE) -->
<!-- } else { -->
<!--   cat("\\textit{PLACEHOLDER: regional table --- knit report.Rmd.}") -->
<!-- } -->
<!-- ``` -->

<!-- ======================================================================
     8. PARETO CHART (combined, PAHO-merged)
     ====================================================================== -->

## Where Impact Is Concentrated

```{r pareto-combined-paho, fig.width=12, fig.height=5.6, out.width="\\linewidth"}
if (!is.null(p_pd_pareto_combined)) {
  print(p_pd_pareto_combined)
} else {
  plot.new()
  text(0.5, 0.5,
       "PLACEHOLDER: PAHO-merged combined Pareto chart (knit report.Rmd)",
       cex = 1.2)
}
```

\footnotesize  Top 20 entries. All Latin America and Caribbean countries combined into a single "PAHO" bar.

<!-- ======================================================================
     9. GLOBAL COVERAGE MAP
     ====================================================================== -->

## Global Coverage: HTN and TFA Interventions

```{r fig-country-map, fig.width=12, fig.height=5.6, out.width="\\linewidth"}
if (!is.null(fig_country_map)) {
  print(fig_country_map)
} else {
  plot.new()
  text(0.5, 0.5,
       "PLACEHOLDER: country coverage map (knit report.Rmd)",
       cex = 1.2)
}
```

<!-- ======================================================================
     10. TFA POLICY TIMELINE
     ====================================================================== -->

## Trans Fat Elimination: Global Policy Progress

```{r fig-tfa-timeline, fig.width=12, fig.height=5.6, out.width="\\linewidth"}
if (!is.null(fig_tfa_timeline)) {
  print(fig_tfa_timeline)
} else {
  plot.new()
  text(0.5, 0.5,
       "PLACEHOLDER: TFA policy timeline (knit report.Rmd)",
       cex = 1.2)
}
```

<!-- ======================================================================
     11. HTN PROGRAM ENROLLMENT (existing asset, kept)
     ======================================================================  -->

## RTSL Hypertension Program Enrollment Over Time

```{r fig-enrollment-control-trend, out.width="90%"}
if (file.exists(png_enrollment)) {
  knitr::include_graphics(png_enrollment)
} else {
  plot.new()
  text(0.5, 0.5,
       "PLACEHOLDER: enrollment trend (knit report.Rmd)",
       cex = 1.2)
}
```

\footnotesize Enrollment and program-controlled population. Post-2025 values held constant at 2025 observed levels.

<!-- ======================================================================
     12. WHAT WE LEARNED
     ====================================================================== -->

## What We Learned {.fragile}

  <!-- \item The \textbf{population reach} of trans-fat policies rivals the impact of HTN treatment --- with no per-patient delivery cost once a policy is in force. -->
  
\begin{columns}[T]
\begin{column}{0.48\textwidth}
\textbf{\color{uwpurple}What surprised us}
\footnotesize
\begin{itemize}
  \item A small number of countries drive most of the global impact: scale of the program matters as much as its efficacy.
\end{itemize}

\medskip
\textbf{\color{uwpurple}What we would do differently}
\footnotesize
\begin{itemize}
  \item Pair modeling more closely with real-time program data feeds.
  \item Invest earlier in a shared data dictionary across country teams.
\end{itemize}
\end{column}

\begin{column}{0.48\textwidth}
\textbf{\color{uwpurple}Barriers we encountered}
\footnotesize
\begin{itemize}
  \item Uneven and delayed country reporting on BP control and program enrollment.
  \item Heterogeneous TFA policy data (passage vs. enforcement dates).
  \item Baseline CVD epidemiology shifts between GBD releases.
  \item Counterfactual definition is contested --- small methodological choices move totals substantially.
\end{itemize}
\end{column}
\end{columns}

<!-- ======================================================================
     13. WHAT'S NEXT
     ====================================================================== -->

## What's Next {.fragile}

\footnotesize
\begin{itemize}
  \item \textbf{Extend} the modeling framework to additional RTSL priority countries and to secondary risk factors (sodium).
  \item \textbf{Refresh} country estimates on an annual cadence as program and policy data arrive.
  \item \textbf{Translate} findings into a country-facing, interactive tool for ministries of health and partners.
  \item \textbf{Disseminate} results through peer-reviewed publication and global health forums to inform policy and funding decisions. "100 Million Lives Saved" manuscript currently under review.
\end{itemize}

\medskip
<!-- \textbf{\color{uwpurple}Our ask of partners} -->
<!-- \footnotesize -->
<!-- \begin{itemize} -->
<!--   \item Country connections --- introductions to ministries for data validation and dissemination. -->
<!--   \item Shared data --- continued access to program monitoring and policy implementation data. -->
<!--   \item Advocacy alignment --- joint messaging on preventable CVD burden for upcoming global forums. -->
<!-- \end{itemize} -->

<!-- \bigskip -->
\begin{exampleblock}{One ask}
\large Help us turn `r grand_combined_m` million projected lives saved into a shared, actionable evidence base for ministries, donors, and advocates.
\end{exampleblock}

<!-- ======================================================================
     EXTRA SLIDES (appendix) --- keep exactly one methodology slide
     ====================================================================== -->

# Extra Slides

## Methodology at a Glance {.fragile}

\begin{columns}[T]
\begin{column}{0.48\textwidth}
\textbf{\color{uwpurple}Model structure}
\footnotesize
\begin{itemize}
  \item Discrete-time, country-by-country multi-state model (Well $\rightarrow$ Sick $\rightarrow$ Dead), ages 20--95, 2017--2050.
  \item Four CVD causes modeled: IHD, ischemic stroke, intracerebral hemorrhage, hypertensive heart disease.
  \item Baseline rates from GBD 2023 (IHME); population from UNWPP 2024.
\end{itemize}
\end{column}

\begin{column}{0.48\textwidth}
\textbf{\color{uwpurple}Interventions and counterfactual}
\footnotesize
\begin{itemize}
  \item \textbf{HTN:} observed RTSL program control rates (2017--2025) maintained to 2050; Ettehad et al. 2016 treatment effects.
  \item \textbf{TFA:} full elimination of industrial TFA in countries with enacted best-practice policies.
  \item \textbf{Counterfactual:} zero incremental program effect (business-as-usual).
  \item \textbf{Combined:} HTN and TFA filtered separately (different target populations), then summed per country.
\end{itemize}
\end{column}
\end{columns}

\vspace{0.4em}
\footnotesize \textit{Inputs: GBD 2023, GBD 2019 RRs, UNWPP 2024, HTN Global Tracker, WHO TFA policy tracker, RTSL program monitoring data.}

<!-- end of presentation -->
